---
title: "Prediction Model using qPCR data (TROST field) to validate breeder trials"
author: "Heike Sprenger"
date: "Wednesday, March 30, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
---


# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_qpcr/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(reshape)
library(pander)
library(pcaMethods)
library(ggplot2)
library(corrplot)
library(gplots)
library(RColorBrewer)

library(plyr)
library(rpart)
library(randomForest)
library(partykit)
library(party)
library(caret) 
library(pls)
library(BioMark)
library(varSelRF)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("qpcr_model_field_breeder.RData")
```


# Load Data
```{r load data}
tolerance <- read.table("../trost_phenotypes/output/tolerance_phenotyper_metadata_2sub.txt", 
                        sep = "\t", header = TRUE)

# relevel tolerance factors
tolerance$tol_cat2_fve <- factor(tolerance$tol_cat2_fve, levels = c("low", "high"))
tolerance$tol_cat3_fve <- factor(tolerance$tol_cat3_fve, levels = c("low","mid", "high"))

##############

# TROST FIELD data (subset of 43 genes!)
# samplelist_field <- read.table("output/model/samplelist_field.txt", header = T, sep = "\t")
# dim(samplelist_field)

# --> comes from: "qpcr_data_field_and_breeder.Rmd"
log_norm_ct_subset_field <- read.table("output/model/log_norm_ct_field_subset_43_genes.txt", header = T, sep = "\t")
dim(log_norm_ct_subset_field)
# 220 43

##############

# Breeder data
# samplelist_breeder <- read.table("output/breeder/samplelist.txt", header = T, sep = "\t")
# dim(samplelist_breeder)
log_norm_ct_breeder <- read.table("output/breeder/log_norm_ct.txt", header = T, sep = "\t")
dim(log_norm_ct_breeder)
# 203 43

##############

# TROST + Breeder data --> comes from: "qpcr_data_field_and_breeder.Rmd"
samplelist_all <- read.table("output/model/samplelist_field_breeder.txt", header = T, sep = "\t")
log_norm_ct_all <- read.table("output/model/log_norm_ct_field_breeder.txt", header = T, sep = "\t")
dim(samplelist_all)
dim(log_norm_ct_all)
# 423 43

head(samplelist_all)

# samplelist only for TROST FIELD data
samplelist_field <- droplevels(subset(samplelist_all, samplelist_all$trost_breeder == "field"))

# samplelist only for BREEDER data
samplelist_breeder <- droplevels(subset(samplelist_all, samplelist_all$trost_breeder == "breeder"))
```


## Define TROST training data
```{r define training data}
# join metadata
head(tolerance)
head(samplelist_field)

# join TROST FIELD samplelist with tolerance information
samplelist_field_joined <- join(samplelist_field, tolerance, by = "subspecies_id")
head(samplelist_field_joined)
dim(samplelist_field_joined)

# only TROST field data
dim(log_norm_ct_subset_field)

# train/test data
train_data <- subset(log_norm_ct_subset_field, samplelist_field_joined$model_set=="train")
test_data <- subset(log_norm_ct_subset_field, samplelist_field_joined$model_set=="test")

# train/test info
train_info <- subset(samplelist_field_joined, samplelist_field_joined$model_set=="train")
test_info <- subset(samplelist_field_joined, samplelist_field_joined$model_set=="test")
```


## Define control data (and meta data)
```{r define control data}
# data
train_data_control <- droplevels(subset(train_data, train_info$treatment == "control"))
test_data_control <- droplevels(subset(test_data, test_info$treatment == "control"))

# meta data
train_info_control <- droplevels(subset(train_info, train_info$treatment == "control"))
test_info_control <- droplevels(subset(test_info, test_info$treatment == "control"))
```


## Define all TROST data (without 3 cultivars --> model_set=="NA") and subset of all TROST control samples
```{r define all data (control)}
# all TROST samples (without NA --> 3 cultivars)
all_data <- subset(log_norm_ct_subset_field, !samplelist_field_joined$model_set=="NA")
all_info <- subset (samplelist_field_joined, !samplelist_field_joined$model_set=="NA")
dim(all_data)
dim(all_info)

# all TROST control samples (without NA --> 3 cultivars)
all_data_control <- subset(log_norm_ct_subset_field, 
                           samplelist_field_joined$treatment=="control" & !samplelist_field_joined$model_set=="NA")

all_info_control <- subset (samplelist_field_joined, 
                            samplelist_field_joined$treatment=="control" & !samplelist_field_joined$model_set=="NA")
```



# Define all Breeder data (without 3 cultivars --> model_set=="NA")
```{r define all breeder data}
samplelist_breeder_joined <- join(samplelist_breeder, tolerance, by="subspecies_id")
dim(samplelist_breeder_joined)

all_info_breeder <- droplevels(subset (samplelist_breeder_joined, !samplelist_breeder_joined$model_set=="NA"))
dim(all_info_breeder)

# information for separate trials or combinations
all_info_norika_2011 <- droplevels(subset(all_info_breeder, all_info_breeder$trial == "Norika_2011"))
all_info_norika_2012 <- droplevels(subset(all_info_breeder, all_info_breeder$trial == "Norika_2012"))
all_info_petersgroden_2011 <- droplevels(subset(all_info_breeder, all_info_breeder$trial == "Petersgroden_2011"))
all_info_petersgroden_2012 <- droplevels(subset(all_info_breeder, all_info_breeder$trial == "Petersgroden_2012"))
all_info_windeby_2011 <- droplevels(subset(all_info_breeder, all_info_breeder$trial == "Windeby_2011"))
all_info_windeby_2012 <- droplevels(subset(all_info_breeder, all_info_breeder$trial == "Windeby_2012"))

# 2011/2012
all_info_2011 <- subset(all_info_breeder, all_info_breeder$treatment == "2011")
all_info_2012 <- subset(all_info_breeder, all_info_breeder$treatment == "2012")

# locations
all_info_norika <- subset(all_info_breeder, all_info_breeder$cultivation == "Norika")
all_info_petersgroden <- subset(all_info_breeder, all_info_breeder$cultivation == "Petersgroden")
all_info_windeby <- subset(all_info_breeder, all_info_breeder$cultivation == "Windeby")
```



# Load data without NAs from PCA (rnipals, no scaling, completeObs)
* comes from: ``qpcr_data_field_and_breeder.Rmd``

## Load data without NAs: TROST FIELD
```{r data without NAs Trost field}
# un-normalized data
log_norm_ct_woNA_all <- read.table("output/model/log_norm_ct_prep_none_rnipals_completeObs_field_breeder.txt",
                                   header = T, sep = "\t")

# ANOVA-normalized data
# log_norm_ct_woNA_all <- read.table("output/model/log_norm_ct_2_prep_none_rnipals_completeObs_field_breeder.txt",
#                                   header = T, sep = "\t")
dim(log_norm_ct_woNA_all)
# 423 43

#### only TROST FIELD data ####
dim(samplelist_all)
log_norm_ct_trost_field_woNA <- subset(log_norm_ct_woNA_all, 
                                       samplelist_all$trost_breeder == "field")
dim(log_norm_ct_trost_field_woNA)
# 220 43

##### train/test #####
train_data_woNA <- subset(log_norm_ct_trost_field_woNA, samplelist_field_joined$model_set=="train")
test_data_woNA <- subset(log_norm_ct_trost_field_woNA, samplelist_field_joined$model_set=="test")
dim(train_data_woNA)
dim(test_data_woNA)

##### control ####
train_data_control_woNA <- subset(train_data_woNA, train_info$treatment=="control")
test_data_control_woNA <- subset(test_data_woNA, test_info$treatment=="control")

##### all (TROST field) ####
all_data_woNA <- subset(log_norm_ct_trost_field_woNA, !samplelist_field_joined$model_set=="NA")
dim(all_data_woNA)
# 202 43

#### all control samples #####
all_data_control_woNA <- subset(log_norm_ct_trost_field_woNA, 
                                samplelist_field_joined$treatment == "control" & !samplelist_field_joined$model_set == "NA")
```


## Load data without NAs: Breeder
```{r data without NAs breeder}
#### only Breeder data #####
log_norm_ct_breeder_woNA <- subset(log_norm_ct_woNA_all, samplelist_all$trost_breeder == "breeder")
dim(log_norm_ct_breeder_woNA)
# 203 43

# without 3 cultivars --> model_set=="NA"
all_data_breeder_woNA <- subset(log_norm_ct_breeder_woNA, !samplelist_breeder_joined$model_set=="NA")
dim(all_data_breeder_woNA)
# 185 43

dim(all_info_breeder)
head(all_info_breeder)

# information for separate trials or combinations
all_data_woNA_norika_2011 <- subset(all_data_breeder_woNA, all_info_breeder$trial == "Norika_2011")
all_data_woNA_norika_2012 <- subset(all_data_breeder_woNA, all_info_breeder$trial == "Norika_2012")
all_data_woNA_petersgroden_2011 <- subset(all_data_breeder_woNA, all_info_breeder$trial == "Petersgroden_2011")
all_data_woNA_petersgroden_2012 <- subset(all_data_breeder_woNA, all_info_breeder$trial == "Petersgroden_2012")
all_data_woNA_windeby_2011 <- subset(all_data_breeder_woNA, all_info_breeder$trial == "Windeby_2011")
all_data_woNA_windeby_2012 <- subset(all_data_breeder_woNA, all_info_breeder$trial == "Windeby_2012")

# 2011/2012
all_data_woNA_2011 <- subset(all_data_breeder_woNA, all_info_breeder$treatment == "2011")
all_data_woNA_2012 <- subset(all_data_breeder_woNA, all_info_breeder$treatment == "2012")

# locations
all_data_woNA_norika <- subset(all_data_breeder_woNA, all_info_breeder$cultivation == "Norika")
all_data_woNA_petersgroden <- subset(all_data_breeder_woNA, all_info_breeder$cultivation == "Petersgroden")
all_data_woNA_windeby <- subset(all_data_breeder_woNA, all_info_breeder$cultivation == "Windeby")
```


# Random forest regression
## Random forest regression: all data
```{r rf reg all}
input_all_woNA <- data.frame(all_data_woNA, "tol" = all_info$mdrym_fve)

set.seed(1)
rf_all <- randomForest(tol~. , data = input_all_woNA, ntree = 1000)
print(rf_all)

#importance(rf_all)
varImpPlot(rf_all)

plot(all_info$mdrym_fve, rf_all$predicted)
abline(0,1)
cor.test(all_info$mdrym_fve, rf_all$predicted)

plot(test_info$mdrym_fve, predict(rf_all, test_data_woNA))
sqrt(mean((test_info$mdrym_fve - predict(rf_all, as.data.frame(test_data_woNA)))^2))
```


## Random forest regression: predict DRYM for breeder data
```{r rf reg all predict DRYM for breeder data}
rf_all_pred_breeder <- predict(rf_all, all_data_breeder_woNA)

plot(all_info_breeder$mdrym_fve, rf_all_pred_breeder, 
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19,
     xlim = c(-0.06, 0.1), ylim = c(-0.06, 0.1))
abline(0, 1, col = "red")

cor.test(all_info_breeder$mdrym_fve, rf_all_pred_breeder)
sqrt(mean((all_info_breeder$mdrym_fve - rf_all_pred_breeder)^2))

# Norika 2011/2012
rf_all_pred_norika_2011 <- predict(rf_all, all_data_woNA_norika_2011)
rf_all_pred_norika_2012 <- predict(rf_all, all_data_woNA_norika_2012)

plot(all_info_norika_2011$mdrym_fve, rf_all_pred_norika_2011, xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19)
text(all_info_norika_2011$mdrym_fve, rf_all_pred_norika_2011, labels = all_info_norika_2011$cultivar, cex = 0.6)
plot(all_info_norika_2012$mdrym_fve, rf_all_pred_norika_2012, xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19)
text(all_info_norika_2012$mdrym_fve, rf_all_pred_norika_2012, labels = all_info_norika_2012$cultivar, cex = 0.6)
```


# Random forest classification
## Random forest classification: all data with 3 classes
```{r rf cat3 all}
input_cat3_all_woNA <- data.frame(all_data_woNA, "tol"=all_info$tol_cat3_fve)
levels(input_cat3_all_woNA$tol)

set.seed(1)
rf_cat3_all <- randomForest(tol~. , data=input_cat3_all_woNA, ntree=1000)
print(rf_cat3_all)

#write.table(importance(rf_cat3_all), "output/model/importance_field_rf_cat3_all.txt", sep="\t")
varImpPlot(rf_cat3_all)

#table(all_info$tol_cat3_fve, rf_cat3_all$predicted)



# only CONTROL data
input_cat3_all_control_woNA <- data.frame(all_data_control_woNA, "tol"=all_info_control$tol_cat3_fve)
levels(input_cat3_all_control_woNA$tol)

set.seed(1)
rf_cat3_all_control <- randomForest(tol~. , data=input_cat3_all_control_woNA, ntree=1000)
print(rf_cat3_all_control)
```


### Cross-Validation
```{r rf cat3 all CV}
set.seed(13)
rf_cat3_all_cv <- rfcv(all_data_woNA, all_info$tol_cat3_fve, step=0.8)
rf_cat3_all_cv$n.var
with(rf_cat3_all_cv, plot(n.var, error.cv, log="x", type="o", lwd=2))
```


### Variable Selection (vs)
```{r rf cat3 all VarSel}
set.seed(1)
rf_cat3_all_vs <- varSelRF(all_data_woNA, all_info$tol_cat3_fve, ntree = 500, ntreeIterat = 300, vars.drop.frac = 0.2, c.sd=1)
rf_cat3_all_vs
# 18
#write.table(rf_cat3_all_vs$selected.vars, "output/model/field_rf_cat3_all_selected_vars.txt", sep="\t")
#rf_cat3_all_vs$selected.vars
plot(rf_cat3_all_vs, which=1)
plot(rf_cat3_all_vs, which=2)

# indices of selected variables
rf_cat3_all_vs_idx <- which(colnames(all_data_woNA) %in% rf_cat3_all_vs$selected.vars)
```


## Random forest classification: selected variables from all data with 3 classes
```{r rf cat3 all selected variables, fig.width=12}
all_data_woNA_vs <- all_data_woNA[,rf_cat3_all_vs_idx]
dim(all_data_woNA_vs)
# 202 18

input_all_cat3_woNA_vs <- data.frame(all_data_woNA_vs, "tol"=all_info$tol_cat3_fve)
levels(input_all_cat3_woNA_vs$tol)

set.seed(1)
rf_cat3_all_reduced <- randomForest(tol~. , data=input_all_cat3_woNA_vs, ntree=1000)
print(rf_cat3_all_reduced)

#write.table(importance(rf_cat3_all_reduced), "output/model/importance_rf_cat3_all_reduced.txt", sep="\t")
par(mfrow=c(1,2))
varImpPlot(rf_cat3_all_reduced)
varImpPlot(rf_cat3_all)
par(mfrow=c(1,1))
```


## Random forest classification: Predict tolerance class for breeder data
```{r rf cat3 predict tolerance class for breeder data}
dim(log_norm_ct_breeder_woNA)

rf_cat3_all_predicted_drym_breeder <- predict(rf_cat3_all, all_data_breeder_woNA)
summary(rf_cat3_all_predicted_drym_breeder)

confusionMatrix(table(rf_cat3_all_predicted_drym_breeder, all_info_breeder$tol_cat3_fve))

table(all_info_breeder$cultivar, rf_cat3_all_predicted_drym_breeder)
table(all_info_breeder$cultivar, all_info_breeder$tol_cat3_fve)


# only CONTROL data
rf_cat3_all_control_predicted_drym_breeder <- predict(rf_cat3_all_control, all_data_breeder_woNA)
confusionMatrix(table(rf_cat3_all_control_predicted_drym_breeder, all_info_breeder$tol_cat3_fve))

# reduced model
# only CONTROL data
rf_cat3_all_reduced_predicted_drym_breeder <- predict(rf_cat3_all_reduced, all_data_breeder_woNA)
confusionMatrix(table(rf_cat3_all_reduced_predicted_drym_breeder, all_info_breeder$tol_cat3_fve))
```


## Random forest classification: Predict tolerance class for single breeder trials
```{r rf cat3 predict tolerance class for single breeder trials}
table( predict(rf_cat3_all, all_data_woNA_norika_2011), all_info_norika_2011$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_woNA_norika_2012), all_info_norika_2012$tol_cat3_fve)

table( predict(rf_cat3_all, all_data_woNA_petersgroden_2011), all_info_petersgroden_2011$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_woNA_petersgroden_2012), all_info_petersgroden_2012$tol_cat3_fve)

table( predict(rf_cat3_all, all_data_woNA_windeby_2011), all_info_windeby_2011$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_woNA_windeby_2012), all_info_windeby_2012$tol_cat3_fve)

# years
confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_2011), all_info_2011$tol_cat3_fve))
confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_2012), all_info_2012$tol_cat3_fve))

# location
table( predict(rf_cat3_all, all_data_woNA_norika), all_info_norika$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_woNA_petersgroden), all_info_petersgroden$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_woNA_windeby), all_info_windeby$tol_cat3_fve)

# accuracy for single trials
confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_norika_2011), 
                       all_info_norika_2011$tol_cat3_fve))$overall[1]
confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_norika_2012), 
                       all_info_norika_2012$tol_cat3_fve))$overall[1]

confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_petersgroden_2011),
                       all_info_petersgroden_2011$tol_cat3_fve))$overall[1]
confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_petersgroden_2012),
                       all_info_petersgroden_2012$tol_cat3_fve))$overall[1]

confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_windeby_2011), 
                       all_info_windeby_2011$tol_cat3_fve))$overall[1]
confusionMatrix(table( predict(rf_cat3_all, all_data_woNA_windeby_2012), 
                       all_info_windeby_2012$tol_cat3_fve))$overall[1]
```


# Lasso Model
## Lasso: Define data
```{r lasso define data}
# TEST data
test_x <- as.matrix(test_data_woNA)
test_y <- test_info$mdrym_fve
test_y_cat <- test_info$tol_cat3_fve

# use TRAIN data
train_x <- as.matrix(train_data_woNA)
train_y <- train_info$mdrym_fve
train_y_cat <- train_info$tol_cat3_fve

# use ALL data
x <- as.matrix(all_data_woNA)
y <- all_info$mdrym_fve
y_cat <- all_info$tol_cat3_fve
```


## Lasso: Regression model with ALL data
```{r lasso reg model with all data}
set.seed(1)
lasso_all <- glmnet(x, y)

# Cross-validation
set.seed(1)
lasso_cv_all <- cv.glmnet(x, y)
plot(lasso_cv_all, ylim=c(0, 0.005))

# use no specific lambda
lasso_all_pred_test <- predict(lasso_all, new = test_x)

# calculate mean error
lasso_all_pred_test_err <- apply( (lasso_all_pred_test - test_y)^2, 2, mean )
points(log(lasso_all$lambda), lasso_all_pred_test_err, col="blue", pch="*")

# use optimal lamba value
lasso_all_pred_all_lambda <- predict(lasso_cv_all, new = x, s="lambda.min")
lasso_all_pred_test_lambda <- predict(lasso_cv_all, new = test_x, s="lambda.min")

# plot observed vs. predicted DRYM
plot(y, lasso_all_pred_all_lambda)
abline(0,1)
cor.test(y, lasso_all_pred_all_lambda)

sqrt(mean((y - lasso_all_pred_all_lambda)^2))
```


## Lasso Regression: Predict DRYM for breeder data
```{r lasso reg predict DRYM for breeder data}
# use optimal lamba value
lasso_all_pred_breeder <- predict(lasso_cv_all, new = as.matrix(all_data_breeder_woNA), s = "lambda.min")

plot(all_info_breeder$mdrym_fve, lasso_all_pred_breeder, 
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19,
     xlim = c(-0.18, 0.15), ylim = c(-0.18, 0.15))
abline(0, 1, col = "red")

cor.test(all_info_breeder$mdrym_fve, lasso_all_pred_breeder)
sqrt(mean((all_info_breeder$mdrym_fve - lasso_all_pred_breeder)^2))

# Norika 2011/2012
lasso_all_pred_norika_2011 <- predict(lasso_cv_all, new = as.matrix(all_data_woNA_norika_2011), s = "lambda.min")
lasso_all_pred_norika_2012 <- predict(lasso_cv_all, new = as.matrix(all_data_woNA_norika_2012), s = "lambda.min")

plot(all_info_norika_2011$mdrym_fve, lasso_all_pred_norika_2011, xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19)
text(all_info_norika_2011$mdrym_fve, lasso_all_pred_norika_2011, labels = all_info_norika_2011$cultivar, cex = 0.6)
plot(all_info_norika_2012$mdrym_fve, lasso_all_pred_norika_2012, xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19)
text(all_info_norika_2012$mdrym_fve, lasso_all_pred_norika_2012, labels = all_info_norika_2012$cultivar, cex = 0.6)
```


## Lasso: Multinomial model with TRAINING data
A special option for multinomial regression is ``type.multinomial``, which allows the usage of a grouped lasso penalty if ``type.multinomial = "grouped"``. 
This will ensure that the multinomial coefficients for a variable are all in or out together, just like for the multi-response Gaussian.
```{r lasso multinomial model with train data}
set.seed(1)
lasso_cat_train <- glmnet(train_x, train_y_cat, family = "multinomial")

# Cross-validation
set.seed(1)
lasso_cat_cv_train <- cv.glmnet(train_x, train_y_cat, family = "multinomial")
plot(lasso_cat_cv_train)

# use optimal lamba value
lasso_cat_train_pred_tain_lambda <- predict(lasso_cat_cv_train, new = train_x, s = "lambda.min", type = "class")
lasso_cat_train_pred_test_lambda <- predict(lasso_cat_cv_train, new = test_x, s = "lambda.min", type = "class")

# table observed vs. predicted DRYM
table(train_y_cat, lasso_cat_train_pred_tain_lambda)
table(test_y_cat, lasso_cat_train_pred_test_lambda)
```


## Lasso: Multinomial model with ALL data
```{r lasso multinomial model with all data}
set.seed(1)
lasso_cat_all <- glmnet(x, y_cat, family = "multinomial")

# Cross-validation
set.seed(1)
lasso_cat_cv_all <- cv.glmnet(x, y_cat, family = "multinomial")
plot(lasso_cat_cv_all)

# use optimal lamba value
lasso_cat_all_pred_all_lambda <- predict(lasso_cat_cv_all, new = x, s = "lambda.min", type = "class")

# plot observed vs. predicted DRYM
table(y_cat, lasso_cat_all_pred_all_lambda)
```


## Lasso multinomial model: Predict DRYM for breeder data
```{r lasso multinomial predict DRYM for breeder data}

func_pred_new_values <- function(new_x, new_y, pred_model){
  
  pred_y <- predict(pred_model, new = as.matrix(new_x), 
                    s = "lambda.min", type  = "class")
  # convert to factor
  pred_y_f <- factor(pred_y[,1], levels = c("low", "mid", "high"))
  
  # print(table(pred_y_f, new_y))
  print(confusionMatrix(table(pred_y_f, new_y)))
}

# all breeder data
func_pred_new_values(new_x = all_data_breeder_woNA,
                     new_y = all_info_breeder$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

# Norika 2011/12
func_pred_new_values(new_x = all_data_woNA_norika_2011,
                     new_y = all_info_norika_2011$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

func_pred_new_values(new_x = all_data_woNA_norika_2012,
                     new_y = all_info_norika_2012$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

# Petersgroden 2011/12
func_pred_new_values(new_x = all_data_woNA_petersgroden_2011,
                     new_y = all_info_petersgroden_2011$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

func_pred_new_values(new_x = all_data_woNA_petersgroden_2012,
                     new_y = all_info_petersgroden_2012$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

# Windeby 2011/12
func_pred_new_values(new_x = all_data_woNA_windeby_2011,
                     new_y = all_info_windeby_2011$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

func_pred_new_values(new_x = all_data_woNA_windeby_2012,
                     new_y = all_info_windeby_2012$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)
```

# Save workspace and sessionInfo
```{r save workspace}
save.image("qpcr_model_field_breeder.RData")
sessionInfo()
```

