---
title: "Prediction Model using qPCR data (TROST field validation)"
author: "Heike Sprenger"
date: "Thursday, March 24, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

**Prediction Model using qPCR data**

* Field Models were trained on 2 datasets and tested on another independent field trial
* Resulted in 3 combinations


# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_qpcr/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(reshape)
library(pander)
library(pcaMethods)
library(ggplot2)
library(corrplot)
library(gplots)
library(RColorBrewer)

library(plyr)
library(rpart)
library(randomForest)
library(partykit)
library(party)
library(caret) 
library(pls)
library(BioMark)
library(varSelRF)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis", tidy = TRUE)
opts_knit$set(root.dir = '../')

# load workspace
#load("qpcr_model_field_cross.RData")
```


# Define samplelist
```{r define samplelist}
tolerance <- read.table("../trost_phenotypes/output/tolerance_phenotyper_metadata_2sub.txt", 
                        sep = "\t", header = TRUE)

# relevel tolerance factors
tolerance$tol_cat2_fve <- factor(tolerance$tol_cat2_fve, levels = c("low", "high"))
tolerance$tol_cat3_fve <- factor(tolerance$tol_cat3_fve, levels = c("low","mid", "high"))

samplelist <- read.table("output/trost/samplelist.txt", header=TRUE, sep="\t")

# join metadata
head(tolerance)
head(samplelist)

# join samplelist
samplelist_joined <- join(samplelist, tolerance, by = "subspecies_id")
samplelist_field <- droplevels(subset(samplelist_joined, samplelist_joined$cultivation == "field"))
```


# Load data without NAs from PCA (rnipals, no scaling)
```{r load data without NAs}
log_norm_ct_woNA <- read.table("output/log_norm_ct_none_rnipals_completeObs.txt", header=TRUE, sep="\t")
dim(log_norm_ct_woNA)

# only field data
log_norm_ct_field_woNA <- subset(log_norm_ct_woNA, samplelist$cultivation=="field")
dim(log_norm_ct_field_woNA)

##### all #### with 31 cultivars
all_data_woNA <- subset(log_norm_ct_field_woNA, !samplelist_field$model_set=="NA")
all_info <- droplevels(subset (samplelist_field, !samplelist_field$model_set=="NA"))
```


## Training data
```{r training data}
# training: two trials
mpi2011_jki2012_data <- subset (all_data_woNA, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="jki_feld_2012")
mpi2011_jki2012_info <- subset (all_info, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="jki_feld_2012")

mpi2011_mpi2012_data <- subset (all_data_woNA, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="mpi_feld_2012")
mpi2011_mpi2012_info <- subset (all_info, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="mpi_feld_2012")

mpi2012_jki2012_data <- subset (all_data_woNA, 
                                all_info$trial=="mpi_feld_2012" | all_info$trial=="jki_feld_2012")
mpi2012_jki2012_info <- subset (all_info, 
                                all_info$trial=="mpi_feld_2012" | all_info$trial=="jki_feld_2012")
```


## Test data
```{r test data}
# test: single trial
mpi2011_data <- subset (all_data_woNA,  all_info$trial=="mpi_feld_2011")
mpi2011_info <- subset (all_info, all_info$trial=="mpi_feld_2011")

mpi2012_data <- subset (all_data_woNA, all_info$trial=="mpi_feld_2012")
mpi2012_info <- subset (all_info, all_info$trial=="mpi_feld_2012")

jki2012_data <- subset (all_data_woNA, all_info$trial=="jki_feld_2012")
jki2012_info <- subset (all_info, all_info$trial=="jki_feld_2012")
```


# Random forest classification
## train data: mpi2011, jki2012 --> test data: mpi2012
```{r rf cat3 mpi2011_jki2012}
input_cat3_mpi2011_jki2012 <- data.frame(mpi2011_jki2012_data, "tol"=mpi2011_jki2012_info$tol_cat3_fve)

set.seed(2)
rf_cat3_mpi2011_jki2012 <- randomForest(tol~. , data=input_cat3_mpi2011_jki2012, ntree=1000)
print(rf_cat3_mpi2011_jki2012)

#importance(rf_cat3_mpi2011_jki2012)
varImpPlot(rf_cat3_mpi2011_jki2012)

table(mpi2011_jki2012_info$tol_cat3_fve, rf_cat3_mpi2011_jki2012$predicted)

table(predict(rf_cat3_mpi2011_jki2012, mpi2012_data), mpi2012_info$tol_cat3_fve)
#       low mid high
#  low   22   3    1
#  mid    0  15    0
#  high   2   6   21

confusionMatrix(table(predict(rf_cat3_mpi2011_jki2012, mpi2012_data), mpi2012_info$tol_cat3_fve))
```


## train data: mpi2011, mpi2012 --> test data: jki2012
```{r rf cat3 mpi2011_mpi2012}
input_cat3_mpi2011_mpi2012 <- data.frame(mpi2011_mpi2012_data, "tol"=mpi2011_mpi2012_info$tol_cat3_fve)

set.seed(2)
rf_cat3_mpi2011_mpi2012 <- randomForest(tol~. , data=input_cat3_mpi2011_mpi2012, ntree=1000)
print(rf_cat3_mpi2011_mpi2012)

#importance(rf_cat3_mpi2011_mpi2012)
varImpPlot(rf_cat3_mpi2011_mpi2012)

table(mpi2011_mpi2012_info$tol_cat3_fve, rf_cat3_mpi2011_mpi2012$predicted)

table(predict(rf_cat3_mpi2011_mpi2012, jki2012_data), jki2012_info$tol_cat3_fve)
#       low mid high
#  low   17   0    1
#  mid    3  22    6
#  high   0   0   13
confusionMatrix(table(predict(rf_cat3_mpi2011_mpi2012, jki2012_data), jki2012_info$tol_cat3_fve))
```


## train data: jki2012, mpi2012 --> test data: mpi2011
```{r rf cat3 mpi2012_jki2012}
input_cat3_mpi2012_jki2012 <- data.frame(mpi2012_jki2012_data, "tol"=mpi2012_jki2012_info$tol_cat3_fve)

set.seed(2)
rf_cat3_mpi2012_jki2012 <- randomForest(tol~. , data=input_cat3_mpi2012_jki2012, ntree=1000)
print(rf_cat3_mpi2012_jki2012)

#importance(rf_cat3_mpi2012_jki2012)
varImpPlot(rf_cat3_mpi2012_jki2012)

table(mpi2012_jki2012_info$tol_cat3_fve, rf_cat3_mpi2012_jki2012$predicted)

table(predict(rf_cat3_mpi2012_jki2012, mpi2011_data), mpi2011_info$tol_cat3_fve)
#       low mid high
#  low   22   0    0
#  mid    1  22    2
#  high   1   2   20
confusionMatrix(table(predict(rf_cat3_mpi2012_jki2012, mpi2011_data), mpi2011_info$tol_cat3_fve))
```



# Random forest regression
## train data: mpi2011, jki2012 --> test data: mpi2012
```{r rf reg mpi2011_jki2012}
input_reg_mpi2011_jki2012 <- data.frame(mpi2011_jki2012_data, "tol"=mpi2011_jki2012_info$mdrym_fve)

rf_reg_mpi2011_jki2012 <- randomForest(tol~. , data=input_reg_mpi2011_jki2012, ntree=1000)
print(rf_reg_mpi2011_jki2012)

#importance(rf_reg_mpi2011_jki2012)
varImpPlot(rf_reg_mpi2011_jki2012)

plot(rf_reg_mpi2011_jki2012$predicted, mpi2011_jki2012_info$mdrym_fve, xlab="predicted", ylab="observed")

plot(predict(rf_reg_mpi2011_jki2012, mpi2012_data), mpi2012_info$mdrym_fve, xlab="predicted", ylab="observed", pch=19)
sqrt(mean((mpi2012_info$mdrym_fve - predict(rf_reg_mpi2011_jki2012, mpi2012_data))^2))
```


## train data: mpi2011, mpi2012 --> test data: jki2012
```{r rf reg mpi2011_mpi2012}
input_reg_mpi2011_mpi2012 <- data.frame(mpi2011_mpi2012_data, "tol"=mpi2011_mpi2012_info$mdrym_fve)

rf_reg_mpi2011_mpi2012 <- randomForest(tol~. , data=input_reg_mpi2011_mpi2012, ntree=1000)
print(rf_reg_mpi2011_mpi2012)

#importance(rf_reg_mpi2011_mpi2012)
varImpPlot(rf_reg_mpi2011_mpi2012)

plot(mpi2011_mpi2012_info$mdrym_fve, rf_reg_mpi2011_mpi2012$predicted)

plot(predict(rf_reg_mpi2011_mpi2012, jki2012_data), jki2012_info$mdrym_fve, xlab="predicted", ylab="observed", pch=19)
sqrt(mean((jki2012_info$mdrym_fve - predict(rf_reg_mpi2011_mpi2012, jki2012_data))^2))
```


## train data: jki2012, mpi2012 --> test data: mpi2011
```{r rf reg mpi2012_jki2012}
input_reg_mpi2012_jki2012 <- data.frame(mpi2012_jki2012_data, "tol"=mpi2012_jki2012_info$mdrym_fve)

rf_reg_mpi2012_jki2012 <- randomForest(tol~. , data=input_reg_mpi2012_jki2012, ntree=1000)
print(rf_reg_mpi2012_jki2012)

#importance(rf_reg_mpi2012_jki2012)
varImpPlot(rf_reg_mpi2012_jki2012)

plot(mpi2012_jki2012_info$mdrym_fve, rf_reg_mpi2012_jki2012$predicted)

plot(predict(rf_reg_mpi2012_jki2012, mpi2011_data), mpi2011_info$mdrym_fve, 
     xlab="predicted", ylab="observed", pch=19, 
     xlim=c(min(mpi2011_info$mdrym_fve), max(mpi2011_info$mdrym_fve)), 
     ylim=c(min(mpi2011_info$mdrym_fve), max(mpi2011_info$mdrym_fve)))
abline(0,1, col="red")

sqrt(mean((mpi2011_info$mdrym_fve - predict(rf_reg_mpi2012_jki2012, mpi2011_data))^2))
```


# Save workspace and sessionInfo
```{r save workspace}
save.image("qpcr_model_field_cross.RData")
sessionInfo()
```

