Prediction Model using qPCR data
Field Models were trained on 2 datasets and tested on another independent field trial
Resulted in 3 combinations
========================================================

```{r setup}
setwd("TROST/Sequencing/qpcr_data/scripts/")
load("qpcr_model_field_cross.RData")
```

### Load packages
```{r load packages}
library("rpart")
library("randomForest")
library("partykit")
library("party")
library(caret) 
library(pls)
library(BioMark)
library(plyr)
```

### Define samplelist
```{r define samplelist}
tolerance <- read.table("../../../Prediction_Model/output/tolerance_phenotyper_metadata_2sub.txt", sep="\t", header=TRUE)
samplelist <- read.table("../output/samplelist.txt", header=TRUE, sep="\t")

tolerance$tol_cat2_fve <- factor(tolerance$tol_cat2_fve, levels = c("low", "high"))
tolerance$tol_cat3_fve <- factor(tolerance$tol_cat3_fve, levels = c("low", "mid", "high"))

# join metadata
head(tolerance)
head(samplelist)

# join samplelist
samplelist_joined <- join(samplelist, tolerance, by="subspecies_id")
samplelist_field <- subset(samplelist_joined, samplelist_joined$cultivation=="field")
```


### data without NAs from PCA (rnipals, no scaling)
```{r data without NAs}
log_norm_ct_woNA <- read.table("../output/log_norm_ct_prep_none_rnipals_completeObs.txt", header=TRUE, sep="\t")
dim(log_norm_ct_woNA)
log_norm_ct_field_woNA <- subset(log_norm_ct_woNA, samplelist$cultivation=="field")
dim(log_norm_ct_field_woNA)

##### all #### with 31 cultivars
all_data_woNA <- subset(log_norm_ct_field_woNA, !samplelist_field$model_set=="NA")
all_info <- subset (samplelist_field, !samplelist_field$model_set=="NA")
```

#### training data
```{r training data}
# training: two trials
mpi2011_jki2012_data <- subset (all_data_woNA, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="jki_feld_2012")
mpi2011_jki2012_info <- subset (all_info, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="jki_feld_2012")

mpi2011_mpi2012_data <- subset (all_data_woNA, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="mpi_feld_2012")
mpi2011_mpi2012_info <- subset (all_info, 
                                all_info$trial=="mpi_feld_2011" | all_info$trial=="mpi_feld_2012")

mpi2012_jki2012_data <- subset (all_data_woNA, 
                                all_info$trial=="mpi_feld_2012" | all_info$trial=="jki_feld_2012")
mpi2012_jki2012_info <- subset (all_info, 
                                all_info$trial=="mpi_feld_2012" | all_info$trial=="jki_feld_2012")
```

#### test data
```{r test data}
# test: single trial
mpi2011_data <- subset (all_data_woNA,  all_info$trial=="mpi_feld_2011")
mpi2011_info <- subset (all_info, all_info$trial=="mpi_feld_2011")

mpi2012_data <- subset (all_data_woNA, all_info$trial=="mpi_feld_2012")
mpi2012_info <- subset (all_info, all_info$trial=="mpi_feld_2012")

jki2012_data <- subset (all_data_woNA, all_info$trial=="jki_feld_2012")
jki2012_info <- subset (all_info, all_info$trial=="jki_feld_2012")
```


### random forest classification: train data
#### train data: mpi2011, jki2012
#### test data: mpi2012
```{r rf cat3 mpi2011_jki2012}
input_cat3_mpi2011_jki2012 <- as.data.frame( cbind(mpi2011_jki2012_data, 
                                              "tol"=mpi2011_jki2012_info$tol_cat3_fve))

set.seed(2)
rf_cat3_mpi2011_jki2012 <- randomForest(tol~. , data=input_cat3_mpi2011_jki2012, ntree=1000)
print(rf_cat3_mpi2011_jki2012)

#importance(rf_cat3_mpi2011_jki2012)
varImpPlot(rf_cat3_mpi2011_jki2012)

table(mpi2011_jki2012_info$tol_cat3_fve, rf_cat3_mpi2011_jki2012$predicted)

table(predict(rf_cat3_mpi2011_jki2012, mpi2012_data), mpi2012_info$tol_cat3_fve)
#       low mid high
#  low   22   3    1
#  mid    0  15    0
#  high   2   6   21

confusionMatrix(table(predict(rf_cat3_mpi2011_jki2012, mpi2012_data), mpi2012_info$tol_cat3_fve))
```


#### train data: mpi2011, mpi2012
#### test data: jki2012
```{r rf cat3 mpi2011_mpi2012}
input_cat3_mpi2011_mpi2012 <- as.data.frame( cbind(mpi2011_mpi2012_data, 
                                              "tol"=mpi2011_mpi2012_info$tol_cat3_fve))

set.seed(2)
rf_cat3_mpi2011_mpi2012 <- randomForest(tol~. , data=input_cat3_mpi2011_mpi2012, ntree=1000)
print(rf_cat3_mpi2011_mpi2012)

#importance(rf_cat3_mpi2011_mpi2012)
varImpPlot(rf_cat3_mpi2011_mpi2012)

table(mpi2011_mpi2012_info$tol_cat3_fve, rf_cat3_mpi2011_mpi2012$predicted)

table(predict(rf_cat3_mpi2011_mpi2012, jki2012_data), jki2012_info$tol_cat3_fve)
#       low mid high
#  low   17   0    1
#  mid    3  22    6
#  high   0   0   13
confusionMatrix(table(predict(rf_cat3_mpi2011_mpi2012, jki2012_data), jki2012_info$tol_cat3_fve))
```


#### train data: jki2012, mpi2012
#### test data: mpi2011
```{r rf cat3 mpi2012_jki2012}
input_cat3_mpi2012_jki2012 <- as.data.frame( cbind(mpi2012_jki2012_data, 
                                              "tol"=mpi2012_jki2012_info$tol_cat3_fve))

set.seed(2)
rf_cat3_mpi2012_jki2012 <- randomForest(tol~. , data=input_cat3_mpi2012_jki2012, ntree=1000)
print(rf_cat3_mpi2012_jki2012)

#importance(rf_cat3_mpi2012_jki2012)
varImpPlot(rf_cat3_mpi2012_jki2012)

table(mpi2012_jki2012_info$tol_cat3_fve, rf_cat3_mpi2012_jki2012$predicted)

table(predict(rf_cat3_mpi2012_jki2012, mpi2011_data), mpi2011_info$tol_cat3_fve)
#       low mid high
#  low   22   0    0
#  mid    1  22    2
#  high   1   2   20
confusionMatrix(table(predict(rf_cat3_mpi2012_jki2012, mpi2011_data), mpi2011_info$tol_cat3_fve))
```



### random forest regression
#### train data: mpi2011, jki2012
#### test data: mpi2012
```{r rf reg mpi2011_jki2012}
input_reg_mpi2011_jki2012 <- as.data.frame( cbind(mpi2011_jki2012_data, 
                                              "tol"=mpi2011_jki2012_info$mdrym_fve))

rf_reg_mpi2011_jki2012 <- randomForest(tol~. , data=input_reg_mpi2011_jki2012, ntree=1000)
print(rf_reg_mpi2011_jki2012)

#importance(rf_reg_mpi2011_jki2012)
varImpPlot(rf_reg_mpi2011_jki2012)

plot(rf_reg_mpi2011_jki2012$predicted, mpi2011_jki2012_info$mdrym_fve, xlab="predicted", ylab="observed")

plot(predict(rf_reg_mpi2011_jki2012, mpi2012_data), mpi2012_info$mdrym_fve, xlab="predicted", ylab="observed", pch=19)
sqrt(mean((mpi2012_info$mdrym_fve - predict(rf_reg_mpi2011_jki2012, mpi2012_data))^2))
```


#### train data: mpi2011, mpi2012
#### test data: jki2012
```{r rf reg mpi2011_mpi2012}
input_reg_mpi2011_mpi2012 <- as.data.frame( cbind(mpi2011_mpi2012_data, 
                                              "tol"=mpi2011_mpi2012_info$mdrym_fve))

rf_reg_mpi2011_mpi2012 <- randomForest(tol~. , data=input_reg_mpi2011_mpi2012, ntree=1000)
print(rf_reg_mpi2011_mpi2012)

#importance(rf_reg_mpi2011_mpi2012)
varImpPlot(rf_reg_mpi2011_mpi2012)

plot(mpi2011_mpi2012_info$mdrym_fve, rf_reg_mpi2011_mpi2012$predicted)

plot(predict(rf_reg_mpi2011_mpi2012, jki2012_data), jki2012_info$mdrym_fve, xlab="predicted", ylab="observed", pch=19)
sqrt(mean((jki2012_info$mdrym_fve - predict(rf_reg_mpi2011_mpi2012, jki2012_data))^2))
```


#### train data: jki2012, mpi2012
#### test data: mpi2011
```{r rf reg mpi2012_jki2012}
input_reg_mpi2012_jki2012 <- as.data.frame( cbind(mpi2012_jki2012_data, 
                                              "tol"=mpi2012_jki2012_info$mdrym_fve))

rf_reg_mpi2012_jki2012 <- randomForest(tol~. , data=input_reg_mpi2012_jki2012, ntree=1000)
print(rf_reg_mpi2012_jki2012)

#importance(rf_reg_mpi2012_jki2012)
varImpPlot(rf_reg_mpi2012_jki2012)

plot(mpi2012_jki2012_info$mdrym_fve, rf_reg_mpi2012_jki2012$predicted)

plot(predict(rf_reg_mpi2012_jki2012, mpi2011_data), mpi2011_info$mdrym_fve, 
     xlab="predicted", ylab="observed", pch=19, 
     xlim=c(min(mpi2011_info$mdrym_fve), max(mpi2011_info$mdrym_fve)), 
     ylim=c(min(mpi2011_info$mdrym_fve), max(mpi2011_info$mdrym_fve)))
abline(0,1, col="red")

sqrt(mean((mpi2011_info$mdrym_fve - predict(rf_reg_mpi2012_jki2012, mpi2011_data))^2))
```



```{r save workspace}
save.image("qpcr_model_field_cross.RData")
```

