---
title: "Analysis of qRT-PCR data from VALDIS trials 2014"
author: "Heike Sprenger"
date: "Thursday, March 24, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_qpcr/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(reshape)
library(pander)
library(pcaMethods)
library(ggplot2)
library(corrplot)
library(gplots)
library(RColorBrewer)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("qpcr_data_valdis.RData")
```


# Source R functions
```{r source R functions}
source("../functions/colors.R")
source("../functions/func_histograms.R")
source("../functions/func_prep_pca.R")
source("../functions/func_pca_plots.R")
```


# Load sample information and modify them
```{r load sample information}
samplelist <- read.table("input/valdis/samplelist.txt", header=TRUE, sep="\t")

# remove 5 samples that could not be measured: 920897, 920922, 922294, 922503, 922554
samplelist <- samplelist[-which(samplelist$sampleID %in% c(920897, 920922, 922294, 922503, 922554)), ]

# order samlelist by ID
samplelist <- samplelist[order(samplelist$sampleID),]

# rename levels of trials
levels(samplelist$trial) <- c("JKI Shelter 2014", "MPI FGH 2014")

table(samplelist$machine, samplelist$trial)
samplelist$machine <- droplevels(samplelist$machine)
samplelist$sector <- droplevels(samplelist$sector)
table(samplelist$sector)

# # add factor cultivation to samplelist
# cultivation <- rep("field", 356)
# cultivation[which(samplelist$trial %in% c("mpi_pruef1", "mpi_pruef3"))] <- "pot"
# samplelist <- cbind(samplelist, cultivation)
# levels(samplelist$cultivation)

# samplelist$subspecies_id2 <- as.factor(samplelist$subspecies_id)
# levels(samplelist$subspecies_id2)[1:4] <- c("Desiree", "Alegria", "Milva", "Saturna")

write.table(samplelist, "output/valdis/samplelist.txt", sep="\t")

# additional column for crossing (general factor for parents)
samplelist$crossing2 <- samplelist$crossing
samplelist$crossing[which(samplelist$crossing %in% c("Ramses", "Euroresa", "Albatros"))] <- "Ramses"
samplelist$crossing <- droplevels(samplelist$crossing)
levels(samplelist$crossing)[4] <- "parents"

# subset of samplelists for trials
samplelist_mpi <- droplevels(subset(samplelist, samplelist$trial=="MPI FGH 2014"))
samplelist_jki <- droplevels(subset(samplelist, samplelist$trial=="JKI Shelter 2014"))
```


# Load qpcr raw data (after cleaning) and modify them
```{r load data}
rawdata <- read.table("input/valdis/rawdata.txt", header=TRUE, sep="\t")

rawdata <- rawdata[,-c(1,3)]
head(rawdata)
colnames(rawdata) <- c("gene_id", "value", "sample_id")

class(rawdata$value)
summary(rawdata$value)

rawdata_cast <- cast(rawdata, sample_id ~ gene_id, value="value")
dim(rawdata_cast)
# rows: 803 samples
# columns 48 genes

rawdata_ct <- rawdata_cast[,2:48]
dim(rawdata_ct)
rownames(rawdata_ct) <- rawdata_cast$sample_id

# order by DMT id
rawdata_ct <- rawdata_ct[, order(colnames(rawdata_ct))]

# values without HK genes 
HK_index <- which(colnames(rawdata_ct) %in% c("PGSC0003DMT400030611", 
                                              "PGSC0003DMT400068117", 
                                              "PGSC0003DMT400037585", 
                                              "PGSC0003DMT400080569"))

rawdata_ct_woHK <-rawdata_ct[,-HK_index]
dim(rawdata_ct_woHK)
# 803 samples, 43 genes
```


# Plot ct values
```{r plot ct values, fig.width=12}
# boxplot of all ct-values vs. genes, HK genes are column 15,22,36,42 (ordered by ID)
color <- rep("grey", 47)
color[c(15,22,36,42)] <- "red"
boxplot(rawdata_ct, las=2, col=color)

# boxplot only for HKG
boxplot(rawdata_ct[,c(15,22,36,42)], las=2)

# plot of HK gene 1 vs. HK gene 2
pairs(rawdata_ct[,c(15,22,36,42)])
plot(rawdata_ct[,36], rawdata_ct[,15])
#identify(rawdata_ct[,36], rawdata_ct[,15])
rawdata_ct[639,]

plot(rawdata_ct[,22], rawdata_ct[,36])
#identify(rawdata_ct[,22], rawdata_ct[,36])
# 50
rawdata_ct[50,]

# replace outlier for sample 920913 (line 50) by NA regarding HKG9 (PGSC0003DMT400068117, column 36)
rawdata_ct[50,36] <- NA

# histogram of ct-values per gene
pdf("figures/valdis/hist_rawdata_ct.pdf")
for (i in 1:43){
hist(rawdata_ct_woHK[,i], breaks = 30, col="grey", 
     main = colnames(rawdata_ct_woHK)[i])
}
dev.off()

# pairs plot of first 10 genes
pairs(rawdata_ct_woHK[,1:10])
```


# Histograms
## Histograms: both trials
```{r histograms: both trials}
rawdata_ct_woHK_2 <- as.data.frame(rawdata_ct_woHK)
rawdata_ct_woHK_2$trial <- samplelist$trial
rawdata_ct_woHK_2$treatment <- samplelist$treatment
rawdata_ct_woHK_2$crossing <- samplelist$crossing2

rawdata_ct_woHK_2$id <- rownames(rawdata_ct_woHK)
#rownames(rawdata_ct_woHK_2) <- NULL
rawdata_ct_woHK_melt <- melt(rawdata_ct_woHK_2)
rawdata_ct_woHK_melt <- rawdata_ct_woHK_melt[order(rawdata_ct_woHK_melt$variable),]

# TREATMENT
pdf("figures/valdis/rawdata_ct_woHK_hist_treatment.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset <- subset(rawdata_ct_woHK_melt, rawdata_ct_woHK_melt$variable == levels(rawdata_ct_woHK_melt$variable)[i])
sub_dataset <- droplevels(sub_dataset)
func_histograms_treatment(sub_dataset) 
}
dev.off()

# TRIAL
pdf("figures/valdis/rawdata_ct_woHK_hist_trial.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset <- subset(rawdata_ct_woHK_melt, rawdata_ct_woHK_melt$variable == levels(rawdata_ct_woHK_melt$variable)[i])
sub_dataset <- droplevels(sub_dataset)
func_histograms_trial(sub_dataset) 
}
dev.off()

# CROSSING
# remove parental line samples!
rawdata_ct_woHK_melt_crossing <- subset(rawdata_ct_woHK_melt, rawdata_ct_woHK_melt$crossing %in% c("AxR", "ExA"))
pdf("figures/valdis/rawdata_ct_woHK_hist_crossing.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset <- subset(rawdata_ct_woHK_melt_crossing, 
                      rawdata_ct_woHK_melt_crossing$variable == levels(rawdata_ct_woHK_melt_crossing$variable)[i])
sub_dataset <- droplevels(sub_dataset)
func_histograms_crossing(sub_dataset) 
}
dev.off()
```


## Histograms: MPI trial
```{r histograms: MPI trial}
rawdata_ct_woHK_mpi <- as.data.frame(rawdata_ct_woHK[1:410,])
rawdata_ct_woHK_mpi$trial <- samplelist_mpi$trial
rawdata_ct_woHK_mpi$treatment <- samplelist_mpi$treatment
rawdata_ct_woHK_mpi$crossing <- samplelist_mpi$crossing2

rawdata_ct_woHK_mpi$id <- rownames(rawdata_ct_woHK_mpi)
rawdata_ct_woHK_mpi_melt <- melt(rawdata_ct_woHK_mpi)
rawdata_ct_woHK_mpi_melt <- rawdata_ct_woHK_mpi_melt[order(rawdata_ct_woHK_mpi_melt$variable),]

source("../functions/func_histograms.R")

# TREATMENT
pdf("figures/valdis/rawdata_ct_woHK_mpi_hist_treatment.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset_mpi <- subset(rawdata_ct_woHK_mpi_melt, 
                          rawdata_ct_woHK_mpi_melt$variable == levels(rawdata_ct_woHK_mpi_melt$variable)[i])
sub_dataset_mpi <- droplevels(sub_dataset_mpi)
func_histograms_treatment(sub_dataset_mpi)
}
dev.off()

# CROSSING
rawdata_ct_woHK_mpi_melt_crossing <- subset(rawdata_ct_woHK_mpi_melt, rawdata_ct_woHK_mpi_melt$crossing %in% c("AxR", "ExA"))
pdf("figures/valdis/rawdata_ct_woHK_mpi_hist_crossing.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset_mpi <- subset(rawdata_ct_woHK_mpi_melt_crossing, 
                          rawdata_ct_woHK_mpi_melt_crossing$variable == levels(rawdata_ct_woHK_mpi_melt_crossing$variable)[i])
sub_dataset_mpi <- droplevels(sub_dataset_mpi)
func_histograms_crossing(sub_dataset_mpi) 
}
dev.off()
```


# Histograms: JKI trial
```{r histograms: JKI trial}
rawdata_ct_woHK_jki <- as.data.frame(rawdata_ct_woHK[411:803,])
rawdata_ct_woHK_jki$trial <- samplelist_jki$trial
rawdata_ct_woHK_jki$treatment <- samplelist_jki$treatment
rawdata_ct_woHK_jki$crossing <- samplelist_jki$crossing

rawdata_ct_woHK_jki$id <- rownames(rawdata_ct_woHK_jki)
rawdata_ct_woHK_jki_melt <- melt(rawdata_ct_woHK_jki)
rawdata_ct_woHK_jki_melt <- rawdata_ct_woHK_jki_melt[order(rawdata_ct_woHK_jki_melt$variable),]

# TREATMENT
pdf("figures/valdis/rawdata_ct_woHK_jki_hist_treatment.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset_jki <- subset(rawdata_ct_woHK_jki_melt, 
                          rawdata_ct_woHK_jki_melt$variable == levels(rawdata_ct_woHK_jki_melt$variable)[i])
sub_dataset_jki <- droplevels(sub_dataset_jki)
func_histograms_treatment(sub_dataset_jki)
}
dev.off()

# CROSSING
rawdata_ct_woHK_jki_melt_crossing <- subset(rawdata_ct_woHK_jki_melt, rawdata_ct_woHK_jki_melt$crossing %in% c("AxR", "ExA"))
pdf("figures/valdis/rawdata_ct_woHK_jki_hist_crossing.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset_jki <- subset(rawdata_ct_woHK_jki_melt_crossing, 
                          rawdata_ct_woHK_jki_melt_crossing$variable == levels(rawdata_ct_woHK_jki_melt_crossing$variable)[i])
sub_dataset_jki <- droplevels(sub_dataset_jki)
func_histograms_crossing(sub_dataset_jki) 
}
dev.off()
```


# Get HK genes per trial
```{r HK genes per trial}
mpi_index <- which(samplelist$trial=="MPI FGH 2014")
jki_index <- which(samplelist$trial=="JKI Shelter 2014")

# MPI
hkg_mpi <- as.data.frame(rawdata_ct[mpi_index, HK_index])
summary(hkg_mpi)
# add columns
hkg_mpi$treatment <- samplelist_mpi$treatment
hkg_mpi$crossing <- samplelist_mpi$crossing2
hkg_mpi$machine <- samplelist_mpi$machine
hkg_mpi$sector <- samplelist_mpi$sector
hkg_mpi$id <- rownames(hkg_mpi)
# melt dataset
hkg_mpi_melt <- melt(hkg_mpi)

# JKI
hkg_jki <- as.data.frame(rawdata_ct[jki_index, HK_index])
summary(hkg_jki)
# add columns
hkg_jki$treatment <- samplelist_jki$treatment
hkg_jki$crossing <- samplelist_jki$crossing
hkg_jki$machine <- samplelist_jki$machine
hkg_jki$sector <- samplelist_jki$sector
hkg_jki$id <- rownames(hkg_jki)
head(hkg_jki)
# melt dataset
hkg_jki_melt <- melt(hkg_jki)
head(hkg_jki_melt)

aggregate(hkg_mpi_melt$value, by=list(hkg_mpi_melt$machine, hkg_mpi_melt$variable), mean, na.rm=T)
aggregate(hkg_jki_melt$value, by=list(hkg_jki_melt$machine, hkg_jki_melt$variable), mean, na.rm=T)

# subset for Machines
hkg_mpi_melt_M1 <- subset(hkg_mpi_melt, hkg_mpi_melt$machine=="M1")
hkg_jki_melt_M1 <- subset(hkg_jki_melt, hkg_jki_melt$machine=="M1")
hkg_mpi_melt_M2 <- subset(hkg_mpi_melt, hkg_mpi_melt$machine=="M2")
hkg_jki_melt_M2 <- subset(hkg_jki_melt, hkg_jki_melt$machine=="M2")
hkg_mpi_melt_M3 <- subset(hkg_mpi_melt, hkg_mpi_melt$machine=="M3")
hkg_jki_melt_M3 <- subset(hkg_jki_melt, hkg_jki_melt$machine=="M3")
hkg_mpi_melt_M4 <- subset(hkg_mpi_melt, hkg_mpi_melt$machine=="M4")
hkg_jki_melt_M4 <- subset(hkg_jki_melt, hkg_jki_melt$machine=="M4")
```


## Plot HK genes per trial
```{r plot HK genes per trial}
pdf("figures/valdis/hkg_boxplots.pdf")
par(mfrow=c(1,2))
par(mar=c(15,4,4,2))

# only MPI
boxplot(rawdata_ct[1:410,c(15,22,36,42)], las=2, main="only MPI samples", ylim=c(19,30), ylab="Ct value")
abline(h=25, col="grey", lwd=2, lty=2)
# only JKI
boxplot(rawdata_ct[411:803,c(15,22,36,42)], las=2, main="only JKI samples", ylim=c(19,30), ylab="Ct value")
abline(h=25, col="grey", lwd=2, lty=2)

# by treatment
palette(cols_treatment)
# only MPI
boxplot(hkg_mpi_melt$value ~ hkg_mpi_melt$treatment * hkg_mpi_melt$variable, 
        col=c(1,2), las=2, main="treatment: only MPI samples", ylab="Ct value", ylim=c(19,30))
# only JKI
boxplot(hkg_jki_melt$value ~ hkg_jki_melt$treatment * hkg_jki_melt$variable, 
        col=c(1,2), las=2, main="treatment: only JKI samples", ylab="Ct value", ylim=c(19,30))
legend("bottomleft", fill=1:2, levels(hkg_mpi_melt$treatment))

par(mfrow=c(1,1))

# by crossing
palette(heike_palette_6)
# only MPI
boxplot(hkg_mpi_melt$value ~ hkg_mpi_melt$crossing * hkg_mpi_melt$variable, 
        col=1:6, las=2, main="crossing: only MPI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:6, levels(hkg_mpi_melt$crossing), horiz = T, cex = 0.8)
# only JKI
boxplot(hkg_jki_melt$value ~ hkg_jki_melt$crossing * hkg_jki_melt$variable, 
        col=1:6, las=2, main="crossing: only JKI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:6, levels(hkg_jki_melt$crossing), horiz = T, cex = 0.8)


# by machine
# only MPI
palette(heike_palette_4)
boxplot(hkg_mpi_melt$value ~ hkg_mpi_melt$machine * hkg_mpi_melt$variable, 
        col=1:4, las=2, main="machine: only MPI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:4, levels(hkg_mpi_melt$machine), horiz = T)
# only JKI
palette(heike_palette_5)
boxplot(hkg_jki_melt$value ~ hkg_jki_melt$machine * hkg_jki_melt$variable, 
        col=1:5, las=2, main="machine: only JKI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:5, levels(hkg_jki_melt$machine), horiz = T)

# by sector
palette(rainbow(8))
# M1
# only MPI
boxplot(hkg_mpi_melt_M1$value ~ hkg_mpi_melt_M1$sector * hkg_mpi_melt_M1$variable, 
        col=1:8, las=2, main="sector (M1): only MPI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_mpi_melt_M1$sector), horiz = T, cex = 0.8)
# only JKI
boxplot(hkg_jki_melt_M1$value ~ hkg_jki_melt_M1$sector * hkg_jki_melt_M1$variable, 
        col=1:8, las=2, main="sector (M1): only JKI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_jki_melt_M1$sector), horiz = T, cex = 0.8)

# M2
# only MPI
boxplot(hkg_mpi_melt_M2$value ~ hkg_mpi_melt_M2$sector * hkg_mpi_melt_M2$variable, 
        col=1:8, las=2, main="sector (M2): only MPI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_mpi_melt_M2$sector), horiz = T, cex = 0.8)
# only JKI
boxplot(hkg_jki_melt_M2$value ~ hkg_jki_melt_M2$sector * hkg_jki_melt_M2$variable, 
        col=1:8, las=2, main="sector (M2): only JKI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_jki_melt_M2$sector), horiz = T, cex = 0.8)

# M3
# only MPI
boxplot(hkg_mpi_melt_M3$value ~ hkg_mpi_melt_M3$sector * hkg_mpi_melt_M3$variable, 
        col=1:8, las=2, main="sector (M3): only MPI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_mpi_melt_M3$sector), horiz = T, cex = 0.8)
# only JKI
boxplot(hkg_jki_melt_M3$value ~ hkg_jki_melt_M3$sector * hkg_jki_melt_M3$variable, 
        col=1:8, las=2, main="sector (M3): only JKI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_jki_melt_M3$sector), horiz = T, cex = 0.8)

# M4
# only MPI
boxplot(hkg_mpi_melt_M4$value ~ hkg_mpi_melt_M4$sector * hkg_mpi_melt_M4$variable, 
        col=1:8, las=2, main="sector (M4): only MPI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_mpi_melt_M4$sector), horiz = T, cex = 0.8)
# only JKI
boxplot(hkg_jki_melt_M4$value ~ hkg_jki_melt_M4$sector * hkg_jki_melt_M4$variable, 
        col=1:8, las=2, main="sector (M4): only JKI samples", ylab="Ct value", ylim=c(19,30))
legend("topright", fill=1:8, levels(hkg_jki_melt_M4$sector), horiz = T, cex = 0.8)

dev.off()

palette("default")
par(mar=c(5,4,4,2))
```


# Calculate mean of 4 housekeeping genes per sample
```{r calculate mean of 4 housekeeping genes per sample}
# mean of 4 housekeeping genes per sample
housekeeping_mean_all <- apply(rawdata_ct[,HK_index], 1, mean, na.rm=TRUE)

# without HKG4 !
# housekeeping_mean <- apply(rawdata_ct[,which(colnames(rawdata_ct) %in% c("PGSC0003DMT400068117", 
#                                                                          "PGSC0003DMT400037585", 
#                                                                          "PGSC0003DMT400080569"))], 
#                            1, mean, na.rm=TRUE)
# plot(housekeeping_mean_all, housekeeping_mean)
```


# Calculate delta Ct: ``CT_gene - mean(CT_HK)``
```{r calculate delta Ct}
delta_ct <- matrix(1, nrow=nrow(rawdata_ct_woHK), ncol=ncol(rawdata_ct_woHK))
colnames(delta_ct) <- colnames(rawdata_ct_woHK)
rownames(delta_ct) <- rownames(rawdata_ct_woHK)

for (i in 1:nrow(rawdata_ct_woHK)){
  for (j in 1:ncol(rawdata_ct_woHK)){
    delta_ct[i,j] <- rawdata_ct_woHK[i,j]-housekeeping_mean_all[i]
    }
  }
```


# Calculate 2 to the power of minus delta Ct
```{r calculate 2 to the power of minus delta Ct}
two_minus_delta_ct <- 2^-(delta_ct)
hist(two_minus_delta_ct, breaks=50, col="grey")
# low values correpond to low expression in comparison to HK genes -> between 0 and 1
# high values correpond to high expression in comparison to HK genes -> larger than 1

# log transform normalized values (2 to the power of minus delta Ct)
log_norm_ct <- log10(two_minus_delta_ct)
hist(log_norm_ct, col="grey", breaks=50)
# low values correpond to low expression in comparison to HK genes (e.g. CT 40)
# high values correpond to high expression in comparison to HK genes (e.g. CT 20)

dim(log_norm_ct)

# remove column 9 due to high number of Ct = 40
log_norm_ct_cleaned <- log_norm_ct[,-9]

write.table(log_norm_ct, "output/valdis/log_norm_ct.txt", sep="\t")
write.table(log_norm_ct_cleaned, "output/valdis/log_norm_ct_cleaned.txt", sep="\t")

log2_norm_ct <- log2(two_minus_delta_ct)
log2_norm_ct[1:5,1:5]
two_minus_delta_ct[1:5,1:5]
```


# PCA
```{r PCA}
log_norm_ct_prep_none_rnipals <- func_prep_pca(log_norm_ct, 
                                               scale_method = "none", 
                                               center_option = FALSE, 
                                               pc_number = 5, 
                                               pca_method = "rnipals")

log_norm_ct_prep_pareto_rnipals <- func_prep_pca(log_norm_ct, 
                                               scale_method = "pareto", 
                                               center_option = FALSE, 
                                               pc_number = 5, 
                                               pca_method = "rnipals")

log_norm_ct_prep_none_ppca <- func_prep_pca(log_norm_ct, 
                                               scale_method = "none", 
                                               center_option = FALSE, 
                                               pc_number = 5, 
                                               pca_method = "ppca")

# export completeObs for later use by RandomForest
write.table(log_norm_ct_prep_none_rnipals@completeObs, 
            "output/valdis/log_norm_ct_prep_none_rnipals_completeObs.txt", sep="\t")

log_norm_ct_prep_pareto_rnipals@R2
log_norm_ct_prep_none_rnipals@R2
```


# PCA plots
```{r pca plots}
pdf("figures/valdis/pca_log_norm_ct.pdf")

# treatment
palette(cols_treatment)
func_5pairs_plot(log_norm_ct_prep_none_rnipals, samplelist, 
                 "treatment", 19, "treatment effect, no scaling")
func_5pairs_plot(log_norm_ct_prep_pareto_rnipals, samplelist, 
                 "treatment", 19, "treatment effect, pareto scaling")
func_pca_plot(log_norm_ct_prep_pareto_rnipals, 1, 2, samplelist, 
              "treatment", 19, 1, "bottomright", 1)
func_pca_plot(log_norm_ct_prep_pareto_rnipals, 1, 3, samplelist, 
              "treatment", 19, 1, "bottomright", 1)

#pairs(log_norm_ct_prep_none_rnipals@scores[,1:3], col=samplelist$treatment, pch=19)
#pairs(log_norm_ct_prep_pareto_rnipals@scores[,1:3], col=samplelist$treatment, pch=19)

# experiment
palette(cols_sample_time2)
func_5pairs_plot(log_norm_ct_prep_none_rnipals, samplelist, "trial", 19, "experiment effect, no scaling")
func_5pairs_plot(log_norm_ct_prep_pareto_rnipals, samplelist, "trial", 19, "experiment effect, pareto scaling")
func_pca_plot(log_norm_ct_prep_pareto_rnipals, 1, 2, samplelist, "trial", 19, 1, "bottomright", 1)

func_pca_plot_sym(log_norm_ct_prep_pareto_rnipals, 1, 2, samplelist, "trial", c(19,17), "treatment", 
                  1, "bottomright", 1, "topright", 1, "trial/treatment (pareto scaling)", levels(samplelist$trial))

# pairs(log_norm_ct_prep_pareto_rnipals@scores[,1:3], col=samplelist$trial, pch=19)
# plot(log_norm_ct_prep_pareto_rnipals@scores[,1], log_norm_ct_prep_pareto_rnipals@scores[,2], col=samplelist$trial, pch=19)
# legend("topright", fill=1:2, c("JKI Shelter 2014", "MPI FGH 2014"))

# pairs(log_norm_ct_prep_pareto_rnipals@scores[,1:3], col=samplelist$trial, pch=c(19,17)[samplelist$treatment])
# plot(log_norm_ct_prep_pareto_rnipals@scores[,1], log_norm_ct_prep_pareto_rnipals@scores[,2], 
#      col=samplelist$trial, pch=c(19,17)[samplelist$treatment])
# legend("topright", fill=1:2, c("JKI Shelter 2014", "MPI FGH 2014"))
# legend("bottomright", pch=c(19,17), legend = levels(samplelist$treatment))

# population
palette(heike_palette_6)
func_5pairs_plot(log_norm_ct_prep_pareto_rnipals, samplelist, "crossing2", 19, "population effect, pareto scaling")
func_pca_plot(log_norm_ct_prep_pareto_rnipals, 1, 3, samplelist, "crossing2", 19, 1.5, "topleft", 0.8)
func_pca_plot(log_norm_ct_prep_pareto_rnipals, 2, 3, samplelist, "crossing2", 19, 1.5, "topleft", 0.8)

# pairs(log_norm_ct_prep_pareto_rnipals@scores[,1:5], col=samplelist$crossing, pch=19)

dev.off()
```


## PCA: only MPI trial
```{r PCA: only MPI trial}
log_norm_ct_mpi <- subset(log_norm_ct, samplelist$trial=="MPI FGH 2014")

log_norm_ct_mpi_prep_none_rnipals <- func_prep_pca(log_norm_ct_mpi, 
                                                   scale_method = "none", 
                                                   center_option = FALSE, 
                                                   pc_number = 5, 
                                                   pca_method = "rnipals")

log_norm_ct_mpi_prep_pareto_rnipals <- func_prep_pca(log_norm_ct_mpi, 
                                                   scale_method = "pareto", 
                                                   center_option = FALSE, 
                                                   pc_number = 5, 
                                                   pca_method = "rnipals")
  
log_norm_ct_mpi_prep_none_ppca <- func_prep_pca(log_norm_ct_mpi, 
                                                   scale_method = "none", 
                                                   center_option = FALSE, 
                                                   pc_number = 5, 
                                                   pca_method = "ppca")


# export completeObs for later use by RandomForest
#write.table(log_norm_ct_prep_none_rnipals@completeObs, "../output/log_norm_ct_prep_none_rnipals_completeObs.txt", sep="\t")
```


## PCA plots: only MPI trial
```{r PCA plots: only MPI trial}
# figures
pdf("figures/valdis/pca_log_norm_ct_mpi.pdf")

# treatment
palette(cols_treatment)
func_5pairs_plot(log_norm_ct_mpi_prep_none_rnipals, samplelist_mpi, "treatment", 19, "treatment effect, no scaling")
func_5pairs_plot(log_norm_ct_mpi_prep_pareto_rnipals, samplelist_mpi, "treatment", 19, "treatment effect, pareto scaling")
func_pca_plot(log_norm_ct_mpi_prep_pareto_rnipals, 1, 2, samplelist_mpi, "treatment", 19, 1, "bottomright", 1)
func_pca_plot(log_norm_ct_mpi_prep_pareto_rnipals, 1, 3, samplelist_mpi, "treatment", 19, 1, "bottomright", 1)
func_pca_plot(log_norm_ct_mpi_prep_pareto_rnipals, 2, 3, samplelist_mpi, "treatment", 19, 1, "bottomright", 1)

# population
palette(heike_palette_6)
func_5pairs_plot(log_norm_ct_mpi_prep_pareto_rnipals, samplelist_mpi, "crossing2", 19, "population effect, pareto scaling")
func_pca_plot(log_norm_ct_mpi_prep_pareto_rnipals, 1, 2, samplelist_mpi, "crossing2", 19, 1.5, "topleft", 0.8)
func_pca_plot(log_norm_ct_mpi_prep_pareto_rnipals, 1, 3, samplelist_mpi, "crossing2", 19, 1.5, "topleft", 0.8)
func_pca_plot(log_norm_ct_mpi_prep_pareto_rnipals, 2, 3, samplelist_mpi, "crossing2", 19, 1.5, "topleft", 0.8)

# population/treatment
func_pca_plot_sym(log_norm_ct_mpi_prep_pareto_rnipals, 1, 2, samplelist_mpi, "crossing2", c(19,17), 
              "treatment", 1.5, "bottomright", 0.8, "topright", 0.8, "population/treatment effect")

dev.off()
```


## PCA: only JKI trial
```{r pca: only JKI trial}
log_norm_ct_jki <- subset(log_norm_ct, samplelist$trial=="JKI Shelter 2014")

log_norm_ct_jki_prep_none_rnipals <- func_prep_pca(log_norm_ct_jki, 
                                                   scale_method = "none", 
                                                   center_option = FALSE, 
                                                   pc_number = 5, 
                                                   pca_method = "rnipals")

log_norm_ct_jki_prep_pareto_rnipals <- func_prep_pca(log_norm_ct_jki, 
                                                   scale_method = "pareto", 
                                                   center_option = FALSE, 
                                                   pc_number = 5, 
                                                   pca_method = "rnipals")
  
log_norm_ct_jki_prep_none_ppca <- func_prep_pca(log_norm_ct_jki, 
                                                   scale_method = "none", 
                                                   center_option = FALSE, 
                                                   pc_number = 5, 
                                                   pca_method = "ppca")
```


## PCA plots: only JKI trial
```{r PCA plots: only JKI trial}
# figures
pdf("figures/valdis/pca_log_norm_ct_jki.pdf")

# treatment
palette(cols_treatment)
func_5pairs_plot(log_norm_ct_jki_prep_none_rnipals, samplelist_jki, "treatment", 19, "treatment effect, no scaling")
func_5pairs_plot(log_norm_ct_jki_prep_pareto_rnipals, samplelist_jki, "treatment", 19, "treatment effect, pareto scaling")
func_pca_plot(log_norm_ct_jki_prep_pareto_rnipals, 1, 2, samplelist_jki, "treatment", 19, 1, "bottomright", 1)
func_pca_plot(log_norm_ct_jki_prep_pareto_rnipals, 1, 3, samplelist_jki, "treatment", 19, 1, "bottomright", 1)
func_pca_plot(log_norm_ct_jki_prep_pareto_rnipals, 2, 3, samplelist_jki, "treatment", 19, 1, "bottomright", 1)

# population
palette(heike_palette_6)
func_5pairs_plot(log_norm_ct_jki_prep_pareto_rnipals, samplelist_jki, "crossing2", 19, "population effect, pareto scaling")
func_pca_plot(log_norm_ct_jki_prep_pareto_rnipals, 1, 2, samplelist_jki, "crossing2", 19, 1.5, "topleft", 0.8)
func_pca_plot(log_norm_ct_jki_prep_pareto_rnipals, 1, 3, samplelist_jki, "crossing2", 19, 1.5, "topleft", 0.8)
func_pca_plot(log_norm_ct_jki_prep_pareto_rnipals, 2, 3, samplelist_jki, "crossing2", 19, 1.5, "topleft", 0.8)

# population/treatment
func_pca_plot_sym(log_norm_ct_jki_prep_pareto_rnipals, 1, 2, samplelist_jki, "crossing2", c(19,17), 
              "treatment", 1.5, "bottomright", 0.8, "topright", 0.8, "population/treatment effect")

dev.off()
```


## PCA plots for REPORT
```{r PCA plots for REPORT}

func_pca_plot(log_norm_ct_prep_pareto_rnipals, 1, 2, samplelist, "treatment", 19, 1, "bottomright", 1)
func_pca_plot(log_norm_ct_prep_pareto_rnipals, 2, 3, samplelist, "treatment", 19, 1, "bottomright", 1)

# WITHOUT DESIREE
samplelist_without_desiree <- subset(samplelist, samplelist$crossing != "Desiree")
pca_scores_without_desiree <- subset(log_norm_ct_prep_pareto_rnipals@scores, samplelist$crossing != "Desiree")
log_norm_ct_prep_pareto_rnipals@R2


pdf("figures/valdis/PCA_scoresplots_report.pdf", width=8, height=6)

theme_new <- theme_set(theme_bw())
theme_new <- theme_update(
  axis.title.x = element_text(face="bold", size=16),
  axis.title.y = element_text(face="bold", size=16, angle=90),
  axis.text.x = element_text(size=12),
  axis.text.y = element_text(size=12),
  legend.title = element_text(face="bold", size=14),
  legend.text = element_text(size=12),
  plot.title = element_text(face="bold", size=16))

# TREATMENT
p <- qplot(log_norm_ct_prep_pareto_rnipals@scores[,1], log_norm_ct_prep_pareto_rnipals@scores[,2], 
           colour = samplelist$treatment, 
           xlab = "PC1 (20.6%)", ylab = "PC2 (11.3%)", 
           size = I(3)) 

p + scale_colour_manual(values = cols_treatment_report, name="Behandlung",
                        labels = c("Kontrolle","Stress"))


p <- qplot(log_norm_ct_prep_pareto_rnipals@scores[,2], log_norm_ct_prep_pareto_rnipals@scores[,3], 
           colour = samplelist$treatment, 
           xlab = "PC2 (11.3%)", ylab = "PC3 (7.1%)", 
           size = I(3)) 

p + scale_colour_manual(values = cols_treatment_report, name="Behandlung",
                        labels = c("Kontrolle","Stress"))

# GENOTYPE
q <- qplot(pca_scores_without_desiree[,2], pca_scores_without_desiree[,3], 
           colour = samplelist_without_desiree$crossing2, 
           xlab = "PC2 (11.3%)", ylab = "PC3 (7.1%)", 
           shape = samplelist_without_desiree$crossing,
           size = I(3)) 

q + scale_colour_manual(values = cols_genotype_report, name="Genotyp",
                        labels = c("Albatros","AxR","Euroresa","ExA","Ramses")) + guides(shape=FALSE)

# EXPERIMENT
p <- qplot(log_norm_ct_prep_pareto_rnipals@scores[,1], log_norm_ct_prep_pareto_rnipals@scores[,2], 
           colour = samplelist$trial, 
           xlab = "PC1 (20.6%)", ylab = "PC2 (11.3%)", 
           size = I(3)) 

p + scale_colour_manual(values = c("darkslateblue", "deeppink4"), name="Experiment",
                        labels = c("JKI Shelter 2015","MPI FGH 2015"))

dev.off()
```


### PCA plots for REPORT MPI
```{r PCA plots for REPORT MPI}
# WITHOUT DESIREE
mpi_samplelist_without_desiree <- subset(samplelist_mpi, samplelist_mpi$crossing != "Desiree")
mpi_pca_scores_without_desiree <- subset(log_norm_ct_mpi_prep_pareto_rnipals@scores, samplelist_mpi$crossing != "Desiree")

log_norm_ct_mpi_prep_pareto_rnipals@R2



pdf("figures/valdis/PCA_scoresplots_report_mpi.pdf", width=8, height=6)

theme_new <- theme_set(theme_bw())
theme_new <- theme_update(
  axis.title.x = element_text(face="bold", size=16),
  axis.title.y = element_text(face="bold", size=16, angle=90),
  axis.text.x = element_text(size=12),
  axis.text.y = element_text(size=12),
  legend.title = element_text(face="bold", size=14),
  legend.text = element_text(size=12),
  plot.title = element_text(face="bold", size=16))

# TREATMENT
p <- qplot(mpi_pca_scores_without_desiree[,1], mpi_pca_scores_without_desiree[,2], 
           colour = mpi_samplelist_without_desiree$treatment, 
           xlab = "PC1 (9.8%)", ylab = "PC2 (7.4%)", 
           size = I(3)) 

p + scale_colour_manual(values = cols_treatment_report, name="Behandlung",
                        labels = c("Kontrolle","Stress"))


# GENOTYPE
q <- qplot(mpi_pca_scores_without_desiree[,2], mpi_pca_scores_without_desiree[,3], 
           colour = mpi_samplelist_without_desiree$crossing2, 
           xlab = "PC2 (7.4%)", ylab = "PC3 (6.9%)", 
           shape = mpi_samplelist_without_desiree$crossing,
           size = I(3)) 

q + scale_colour_manual(values = cols_genotype_report, name="Genotyp",
                        labels = c("Albatros","AxR","Euroresa","ExA","Ramses")) + guides(shape=FALSE)

dev.off()
```


### PCA plots for REPORT JKI
```{r PCA plots for REPORT JKI}
# WITHOUT DESIREE
jki_samplelist_without_desiree <- subset(samplelist_jki, samplelist_jki$crossing != "Desiree")
jki_pca_scores_without_desiree <- subset(log_norm_ct_jki_prep_pareto_rnipals@scores, samplelist_jki$crossing != "Desiree")

log_norm_ct_jki_prep_pareto_rnipals@R2



pdf("figures/valdis/PCA_scoresplots_report_jki.pdf", width=8, height=6)

theme_new <- theme_set(theme_bw())
theme_new <- theme_update(
  axis.title.x = element_text(face="bold", size=16),
  axis.title.y = element_text(face="bold", size=16, angle=90),
  axis.text.x = element_text(size=12),
  axis.text.y = element_text(size=12),
  legend.title = element_text(face="bold", size=14),
  legend.text = element_text(size=12),
  plot.title = element_text(face="bold", size=16))

# TREATMENT
p <- qplot(jki_pca_scores_without_desiree[,1], jki_pca_scores_without_desiree[,2], 
           colour = jki_samplelist_without_desiree$treatment, 
           xlab = "PC1 (28.8%)", ylab = "PC2 (9.3%)", 
           size = I(3)) 

p + scale_colour_manual(values = cols_treatment_report, name="Behandlung",
                        labels = c("Kontrolle","Stress"))


# GENOTYPE
q <- qplot(jki_pca_scores_without_desiree[,2], jki_pca_scores_without_desiree[,3], 
           colour = jki_samplelist_without_desiree$crossing2, 
           xlab = "PC2 (9.3%)", ylab = "PC3 (5.6%)", 
           shape = jki_samplelist_without_desiree$crossing,
           size = I(3)) 

q + scale_colour_manual(values = cols_genotype_report, name="Genotyp",
                        labels = c("Albatros","AxR","Euroresa","ExA","Ramses")) + guides(shape=FALSE)

dev.off()
```


# Boxplots
## Boxplots per gene MPI
```{r boxplots per gene MPI}
log_norm_ct_mpi_2 <- as.data.frame(log_norm_ct_mpi)
log_norm_ct_mpi_2$treatment <- samplelist_mpi$treatment
log_norm_ct_mpi_2$crossing <- samplelist_mpi$crossing2
log_norm_ct_mpi_2$id <- rownames(log_norm_ct_mpi_2)

log_norm_ct_mpi_melt <- melt(log_norm_ct_mpi_2)
log_norm_ct_mpi_melt <- log_norm_ct_mpi_melt[order(log_norm_ct_mpi_melt$variable),]

# TREATMENT
pdf("figures/valdis/boxplot_log_norm_ct_mpi_treatment.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset_mpi <- subset(log_norm_ct_mpi_melt, 
                          log_norm_ct_mpi_melt$variable == levels(log_norm_ct_mpi_melt$variable)[i])
sub_dataset_mpi <- droplevels(sub_dataset_mpi)
boxplot(sub_dataset_mpi$value ~ sub_dataset_mpi$treatment)
}
dev.off()
```


## Boxplots per gene JKI
```{r boxplots per gene JKI}
log_norm_ct_jki_2 <- as.data.frame(log_norm_ct_jki)
log_norm_ct_jki_2$treatment <- samplelist_jki$treatment
log_norm_ct_jki_2$crossing <- samplelist_jki$crossing2
log_norm_ct_jki_2$id <- rownames(log_norm_ct_jki_2)

log_norm_ct_jki_melt <- melt(log_norm_ct_jki_2)
log_norm_ct_jki_melt <- log_norm_ct_jki_melt[order(log_norm_ct_jki_melt$variable),]

# TREATMENT
pdf("figures/valdis/boxplot_log_norm_ct_jki_treatment.pdf", width=6, height=6)
for(i in 1:43) {
sub_dataset_jki <- subset(log_norm_ct_jki_melt, 
                          log_norm_ct_jki_melt$variable == levels(log_norm_ct_jki_melt$variable)[i])
sub_dataset_jki <- droplevels(sub_dataset_jki)
boxplot(sub_dataset_jki$value ~ sub_dataset_jki$treatment)
}
dev.off()
```


# Save workspace
```{r save workspace}
save.image("qpcr_data_valdis.RData")
sessionInfo()
```

