---
title: "Combine qRT-PCR data from TROST FIELD trials and Breeder trials"
author: "Heike Sprenger"
date: "April 26, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_qpcr/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(reshape)
library(pander)
library(pcaMethods)
library(ggplot2)
library(corrplot)
library(gplots)
library(RColorBrewer)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("qpcr_data_field_and_breeder.RData")
```


# Source R functions
```{r source R functions}
source("../functions/colors.R")
source("../functions/names.R")
source("../functions/func_histograms.R")
source("../functions/func_prep_pca.R")
source("../functions/func_pca_plots.R")
source("../functions/RemoveFactors_function.R")
```


# Load sample information
```{r load sample information}
# TROST samplelist
samplelist_trost <- read.table("output/trost/samplelist.txt", header=TRUE, sep="\t")
dim(samplelist_trost)

samplelist_trost_part <- samplelist_trost[, c("cultivation", "trial", 
                                              "sampleID", "aliquotID", "cultivar", 
                                              "subspecies_id", "treatment")]

levels(samplelist_trost_part$treatment) <- c("control", "drought stress")

samplelist_field_part <- droplevels(subset(samplelist_trost_part, 
                                           samplelist_trost_part$cultivation == "field"))
dim(samplelist_field_part)
# 220 7

# save samplelist for trost field
write.table(samplelist_field_part, "output/model/samplelist_field.txt", sep="\t")

# breeder samplelist
samplelist_breeder <- read.table("output/breeder/samplelist.txt", header=TRUE, sep="\t")
samplelist_breeder$year <- as.factor(samplelist_breeder$year)

samplelist_breeder_part <- samplelist_breeder[, c("location_name", "trial", 
                                                "sampleID", "aliquotID", "cultivar", 
                                                "subspecies_id", "year")]


# use year (breeder) as equivalent for treatment (TROST)
# use location_name (breeder) as equivalent for cultivation (TROST)
colnames(samplelist_breeder_part)[c(1,7)] <- c("cultivation", "treatment")


# combine both samplelists
samplelist_all <- rbind(samplelist_field_part, samplelist_breeder_part)
dim(samplelist_all)
# 423 7

# add column for trost field or breeder
samplelist_all$trost_breeder <- as.factor( c(rep("field", 220), rep("breeder", 203)))

# save common samplelist for trost and breeder
write.table(samplelist_all, "output/model/samplelist_field_breeder.txt", sep="\t")
```


# Load list of short cultivar names
```{r load list of short cultivar names}
cultivar_names_short <- read.table("cultivar_short_names.txt", header = T, sep = "\t")

# merge with samplelist
samplelist_all <- merge(samplelist_all, cultivar_names_short, by = "cultivar")
head(samplelist_all)

# order samplelist by sampleID
samplelist_all <- samplelist_all[order(samplelist_all$sampleID),]

# save common samplelist for trost and breeder
write.table(samplelist_all, "output/model/samplelist_field_breeder.txt", sep="\t")
```


# Load data
```{r load data}
# TROST data
log_norm_ct_trost <- read.table("output/trost/log_norm_ct.txt", header=TRUE, sep="\t")
dim(log_norm_ct_trost)
# 356 88

rownames(log_norm_ct_trost) <- samplelist_trost$sampleID

# Breeder data
log_norm_ct_breeder <- read.table("output/breeder/log_norm_ct.txt", header=TRUE, sep="\t")
dim(log_norm_ct_breeder)
# 203 43

# rownames(log_norm_ct_breeder)
```


# Subset of 43 valdis/breeder genes
```{r subset of 43 valdis/breeder genes}
# save colnames in variable
log_norm_ct_breeder_colnames <- colnames(log_norm_ct_breeder)
# DMT

# order dataframe by colnames and save colnames in variable
log_norm_ct_trost_ordered <- log_norm_ct_trost[, order(colnames(log_norm_ct_trost))]
log_norm_ct_trost_colnames <- colnames(log_norm_ct_trost_ordered)
# DMT

# index of 42 genes used for trost AND valdis/breeder
overlapping_genes_idx <- which(log_norm_ct_trost_colnames %in% log_norm_ct_breeder_colnames)
log_norm_ct_subset_trost <- log_norm_ct_trost_ordered[, overlapping_genes_idx]
dim(log_norm_ct_subset_trost)

# only TROST field samples
log_norm_ct_subset_field <- subset(log_norm_ct_subset_trost, 
                                   samplelist_trost_part$cultivation == "field")
write.table(log_norm_ct_subset_field, "output/model/log_norm_ct_field_subset_43_genes.txt", sep = "\t")

dim(log_norm_ct_subset_field)
# 220 43
dim(log_norm_ct_breeder)
# 203 43

# combine both datasets
log_norm_ct_all <- rbind(log_norm_ct_subset_field, log_norm_ct_breeder)
dim(log_norm_ct_all)
# 423 43

# order data by sampleID
log_norm_ct_all <- log_norm_ct_all[order(rownames(log_norm_ct_all)), ]

sum(is.na(log_norm_ct_all))

# save common dataset for trost and valdis
write.table(log_norm_ct_all, "output/model/log_norm_ct_field_breeder.txt", sep = "\t")
```


# Remove Factors by ANOVA models using R script by Jan Lisec
```{r normalize (Remove Factors)}
levels(samplelist_all$trial)
levels(samplelist_all$trost_breeder)
levels(samplelist_all$treatment)
levels(samplelist_all$cultivar)

# with trial and cultivar
log_norm_ct_all_2 <- apply(log_norm_ct_all, 2, RemoveFactors, sam = samplelist_all, 
                       facs=c("cultivar", "trial"), 
                       keep=c("cultivar"))

# # with treatment and cultivar
# log_norm_ct_all_2 <- apply(log_norm_ct_all, 2, RemoveFactors, sam = samplelist_all, 
#                        facs=c("cultivar", "treatment"), 
#                        keep=c("cultivar"))
```


## Heatmap of normalized expression data
```{r Heatmap of normalized expression data}
heatmap.2(log_norm_ct_all_2,
          trace = "none")
```


# PCA
## PCA of un-normalized data
```{r PCA of un-normalized data}

log_norm_ct_none_rnipals <- func_prep_pca(log_norm_ct_all, 
                                          scale_method = "none", 
                                          center_option = FALSE, 
                                          pc_number = 5, pca_method = "rnipals")

# log_norm_ct_pareto_rnipals <- func_prep_pca(log_norm_ct_all, scale_method = "pareto", center_option = FALSE, 
#                                              pc_number = 5, pca_method = "rnipals")

# log_norm_ct_none_ppca <- func_prep_pca(log_norm_ct_all, scale_method = "none", center_option = FALSE, 
#                                              pc_number = 5, pca_method = "ppca")


# export completeObs for later use by RandomForest
write.table(log_norm_ct_none_rnipals@completeObs,
            "output/model/log_norm_ct_prep_none_rnipals_completeObs_field_breeder.txt", sep = "\t")
dim(log_norm_ct_none_rnipals@completeObs)
# 423 samples, 43 genes
```


## PCA plots
```{r pca plots}
pdf("figures/model/pca_log_norm_ct_field_breeder.pdf", width=6, height=6)
par(mar=c(4.3, 4.5, 2, 1))
# no scaling, rnipals
# treatment/year
palette(heike_palette_4)
func_5pairs_plot(log_norm_ct_none_rnipals, samplelist_all, "treatment", 19, "treatment effect")
func_pca_plot(log_norm_ct_none_rnipals, 1, 3, samplelist_all, "treatment", 19, 1, 
              "bottomright", 0.8, maintext="treatment effect")


# trost/breeder samples
palette(cols_sample_time2)
func_5pairs_plot(log_norm_ct_none_rnipals, samplelist_all, "trost_breeder", 19, "trost/breeder effect")
func_pca_plot(log_norm_ct_none_rnipals, 1, 2, samplelist_all, "trost_breeder", 19, 1, 
              "bottomright", 0.8, maintext="trost/breeder effect")


# cultivation/location
palette(heike_palette_4)
func_5pairs_plot(log_norm_ct_none_rnipals, samplelist_all, "cultivation", 19, "cultivation effect")
func_pca_plot(log_norm_ct_none_rnipals, 1, 2, samplelist_all, "cultivation", 19, 1, 
              "bottomright", 0.8, maintext="cultivation effect")

# trial: 11
# palette(cols_trial_ordered)
palette(rainbow(9))
func_5pairs_plot(log_norm_ct_none_rnipals, samplelist_all, "trial", 19, "trial effect")
func_pca_plot(log_norm_ct_none_rnipals, 1, 2, samplelist_all, "trial", 19, 1, 
              "bottomright", 0.8, maintext="trial effect")

# plot(log_norm_ct_prep_none_rnipals@scores[,1], log_norm_ct_prep_none_rnipals@scores[,2], col=samplelist_all$trial, pch=19)
# legend("topleft", fill=1:7, c("JKI field 2012", "MPI field 2011", "MPI field 2012", "MPI-MP GH3", "MPI-MP GH5", "JKI Shelter 2014", "MPI-MP FGH 2014"))

# genotype
palette(rainbow(34))

# 34 distinct colors from: 
# http://withr.me/blog/2014/06/16/generate-distict-colors/
# palette(c("#77FAD9FF", "#FFC7FFFF", "#A8FD8EFF", "#CF800EFF", "#44E3EEFF", "#7F48EFFF", "#3BEB3CFF", "#9A98FDFF", "#FDA292FF", "#EF46D2FF", "#105FD0FF", "#4BC202FF", "#FAB174FF", "#C0C0C0FF", "#6BBBF8FF", "#B002C0FF", "#A8EF43FF", "#D9FFE4FF", "#05A3C4FF", "#E6D230FF", "#EC693DFF", "#24DFAEFF", "#FBF77BFF", "#C35103FF", "#7003C1FF", "#DC1F31FF", "#C21E03FF", "#1502C1FF", "#FFEEBAFF", "#4461EEFF", "#F8D86AFF", "#C48E03FF", "#44EE89FF", "#C0C0C0FF"))

# http://tools.medialab.sciences-po.fr/iwanthue/
palette(c("#D34428", "#7E3E24", "#DD8935", "#A47C3B", "#CFB031", "#7F7563", "#545921", "#A1AD4D", 
          "#88BE30", "#AAAF7E", "#3F7D29", "#46CC33", "#59C763", "#61AC72", "#3B604A", "#5AC7AB", 
          "#74AAAD", "#66ADD8", "#415D77", "#728CDA", "#AF9CBB", "#584D8D", "#7C63D5", "#C67FD4", 
          "#D150E1", "#6A4D57", "#8B357C", "#DA44B4", "#D881AA", "#D54185", "#8E3851", "#CFA295", 
          "#D53F5A", "#D37765"))

func_5pairs_plot(log_norm_ct_none_rnipals, samplelist_all, "cultivar", 19, "genotype effect")
func_pca_plot(log_norm_ct_none_rnipals, 1, 2, samplelist_all, "cultivar", 19, 1, 
              "bottomright", 0.7, maintext="genotype effect")
func_pca_plot(log_norm_ct_none_rnipals, 1, 4, samplelist_all, "cultivar", 19, 1, 
              "bottomright", 0.7, maintext="genotype effect")

plot(log_norm_ct_none_rnipals@scores[,1], log_norm_ct_none_rnipals@scores[,2], 
     col = samplelist_all$cultivar, pch = 19, cex = 0.8, main = "genotype effect",
     xlab = "PC1", ylab = "PC2")
text(log_norm_ct_none_rnipals@scores[,1], log_norm_ct_none_rnipals@scores[,2], 
     labels = samplelist_all$cultivar_short, cex = 0.7)

dev.off()
```


## PCA for check cultivars
```{r PCA for check cultivars}
log_norm_ct_check_cultivars <- subset(log_norm_ct_all, samplelist_all$cultivar %in% check_names)
samplelist_check_cultivars <- droplevels(subset(samplelist_all, samplelist_all$cultivar %in% check_names))

log_norm_ct_none_rnipals_check_cultivars <- func_prep_pca(log_norm_ct_check_cultivars, scale_method = "none", center_option = FALSE, 
                                             pc_number = 5, pca_method = "rnipals")


pdf("figures/model/pca_log_norm_ct_field_breeder_check_cultivars.pdf", width=5, height=5)
par(mar=c(4.3, 4.5, 2, 1))

palette(cols_cultivar_check)
func_5pairs_plot(log_norm_ct_none_rnipals_check_cultivars, samplelist_check_cultivars, "cultivar", 19, "cultivar effect")
func_pca_plot(log_norm_ct_none_rnipals_check_cultivars, 1, 3, samplelist_check_cultivars, "cultivar", 19, 1, 
              "bottomright", 1, maintext="cultivar effect")

func_pca_plot_sym(pca_res = log_norm_ct_none_rnipals_check_cultivars, dim1 = 1, dim2 = 3, 
                  factors = samplelist_check_cultivars, color_factor = "cultivar", 
                  symbols = c(17,19), symbol_factor = "trost_breeder", symbol_size = 1, 
                  pos1 = "bottomright", leg1 = 1, pos2 = "topright", leg2 = 1,
                  maintext = "check cultivars", legend.text1 = check_names)

palette(heike_palette_2)
func_5pairs_plot(log_norm_ct_none_rnipals_check_cultivars, samplelist_check_cultivars, "trost_breeder", 19, "trost/breeder effect")
func_pca_plot(log_norm_ct_none_rnipals_check_cultivars, 1, 2, samplelist_check_cultivars, "trost_breeder", 19, 1, 
              "bottomright", 1, maintext="trost/breeder effect")
dev.off()
```


## PCA of ANOVA-normalized data
```{r PCA of ANOVA-normalized data}
log_norm_ct_2_none_rnipals <- func_prep_pca(log_norm_ct_all_2, 
                                            scale_method = "none", 
                                            center_option = FALSE, 
                                            pc_number = 5, pca_method = "rnipals")

# export completeObs for later use by RandomForest
write.table(log_norm_ct_2_none_rnipals@completeObs,
            "output/model/log_norm_ct_2_prep_none_rnipals_completeObs_field_breeder.txt", sep = "\t")

dim(log_norm_ct_2_none_rnipals@completeObs)
# 423 samples, 43 genes
```


## PCA plots of ANOVA-normalized data
```{r pca plots of ANOVA-normalized data}
pdf("figures/model/pca_log_norm_ct_2_field_breeder.pdf", width=7, height=7)
par(mar=c(4.3, 4.5, 2, 1))
# no scaling, rnipals
# treatment/year
palette(heike_palette_4)
func_5pairs_plot(log_norm_ct_2_none_rnipals, samplelist_all, "treatment", 19, "treatment effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 1, 2, samplelist_all, "treatment", 19, 1, 
              "bottomright", 0.8, maintext="treatment effect")


# trost/breeder samples
palette(cols_sample_time2)
func_5pairs_plot(log_norm_ct_2_none_rnipals, samplelist_all, "trost_breeder", 19, "trost/breeder effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 1, 2, samplelist_all, "trost_breeder", 19, 1, 
              "bottomright", 0.8, maintext="trost/breeder effect")


# cultivation/location
palette(heike_palette_4)
func_5pairs_plot(log_norm_ct_2_none_rnipals, samplelist_all, "cultivation", 19, "cultivation effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 1, 2, samplelist_all, "cultivation", 19, 1, 
              "bottomright", 0.8, maintext="cultivation effect")

# trial: 11
# palette(cols_trial_ordered)
palette(rainbow(9))
func_5pairs_plot(log_norm_ct_2_none_rnipals, samplelist_all, "trial", 19, "trial effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 1, 2, samplelist_all, "trial", 19, 1, 
              "bottomright", 0.8, maintext="trial effect")

# plot(log_norm_ct_prep_none_rnipals@scores[,1], log_norm_ct_prep_none_rnipals@scores[,2], col=samplelist_all$trial, pch=19)
# legend("topleft", fill=1:7, c("JKI field 2012", "MPI field 2011", "MPI field 2012", "MPI-MP GH3", "MPI-MP GH5", "JKI Shelter 2014", "MPI-MP FGH 2014"))

# genotype
# 34 distinct colors from: 
# http://tools.medialab.sciences-po.fr/iwanthue/
palette(c("#D34428", "#7E3E24", "#DD8935", "#A47C3B", "#CFB031", "#7F7563", "#545921", "#A1AD4D", 
          "#88BE30", "#AAAF7E", "#3F7D29", "#46CC33", "#59C763", "#61AC72", "#3B604A", "#5AC7AB", 
          "#74AAAD", "#66ADD8", "#415D77", "#728CDA", "#AF9CBB", "#584D8D", "#7C63D5", "#C67FD4", 
          "#D150E1", "#6A4D57", "#8B357C", "#DA44B4", "#D881AA", "#D54185", "#8E3851", "#CFA295", 
          "#D53F5A", "#D37765"))

func_5pairs_plot(log_norm_ct_2_none_rnipals, samplelist_all, "cultivar", 19, "genotype effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 1, 2, samplelist_all, "cultivar", 19, 1, 
              "bottomright", 0.8, maintext="genotype effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 2, 3, samplelist_all, "cultivar", 19, 1, 
              "bottomright", 0.8, maintext="genotype effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 1, 4, samplelist_all, "cultivar", 19, 1, 
              "bottomright", 0.8, maintext="genotype effect")
func_pca_plot(log_norm_ct_2_none_rnipals, 4, 5, samplelist_all, "cultivar", 19, 1, 
              "bottomright", 0.8, maintext="genotype effect")

plot(log_norm_ct_2_none_rnipals@scores[,1], log_norm_ct_2_none_rnipals@scores[,2], 
     col = samplelist_all$cultivar, pch = 19, cex = 0.8, main = "genotype effect",
     xlab = "PC1", ylab = "PC2")
text(log_norm_ct_2_none_rnipals@scores[,1], log_norm_ct_2_none_rnipals@scores[,2], 
     labels = samplelist_all$cultivar_short, cex = 0.7)

dev.off()
```


## PCA plot for publication
```{r PCA plot for publication}
pdf("figures/model/pca_for_publication_treatment.pdf", width=7, height=7)
par(mar=c(4.3, 4.5, 2, 1))

palette(heike_palette_4)
func_pca_plot(log_norm_ct_2_none_rnipals, 1, 2, samplelist_all, "treatment", 19, 1, 
              "bottomright", 1, maintext="treatment effect", 
              legend.text = c("field: control", "field: drought stress", 
                              "agronomic trials 2011", "agronomic trials 2012"))
dev.off()



pdf("figures/model/pca_for_publication_genotype.pdf", width=7, height=7)
par(mar=c(4.3, 4.5, 2, 1))
# 34 distinct colors from: http://tools.medialab.sciences-po.fr/iwanthue/
palette(c("#D34428", "#7E3E24", "#DD8935", "#A47C3B", "#CFB031", "#7F7563", "#545921", "#A1AD4D", 
          "#88BE30", "#AAAF7E", "#3F7D29", "#46CC33", "#59C763", "#61AC72", "#3B604A", "#5AC7AB", 
          "#74AAAD", "#66ADD8", "#415D77", "#728CDA", "#AF9CBB", "#584D8D", "#7C63D5", "#C67FD4", 
          "#D150E1", "#6A4D57", "#8B357C", "#DA44B4", "#D881AA", "#D54185", "#8E3851", "#CFA295", 
          "#D53F5A", "#D37765"))

func_pca_plot(log_norm_ct_2_none_rnipals, 1, 2, samplelist_all, "cultivar", 19, 1,
              "topleft", 0.8, maintext="genotype effect", xmin = -6)

func_pca_plot(log_norm_ct_2_none_rnipals, 2, 3, samplelist_all, "cultivar", 19, 1,
              "topleft", 0.8, maintext="genotype effect", xmin = -3.3)

func_pca_plot(log_norm_ct_2_none_rnipals, 4, 5, samplelist_all, "cultivar", 19, 1, 
              "topright", 0.8, maintext="genotype effect", xmax = 3.3)
dev.off()
```


# Save workspace and sessioninfo
```{r save workspace}
save.image("qpcr_data_field_and_breeder.RData")
sessionInfo()
```

