---
title: "Prediction Model using qPCR data (TROST greenhouse/pot)"
author: "Heike Sprenger"
date: "Thursday, March 24, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---


# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_qpcr/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(reshape)
library(pander)
library(pcaMethods)
library(ggplot2)
library(corrplot)
library(gplots)
library(RColorBrewer)

library(plyr)
library(rpart)
library(randomForest)
library(partykit)
library(party)
library(caret) 
library(pls)
library(BioMark)
library(varSelRF)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("qpcr_model_pot.RData")
```


# Load Data
```{r load data}
tolerance <- read.table("../trost_phenotypes/output/tolerance_phenotyper_metadata_2sub.txt", 
                        sep = "\t", header = TRUE)

# relevel tolerance factors
tolerance$tol_cat2_gh <- factor(tolerance$tol_cat2_gh, levels = c("low", "high"))
tolerance$tol_cat3_gh <- factor(tolerance$tol_cat3_gh, levels = c("low","mid", "high"))
tolerance$tol_cat2_fve <- factor(tolerance$tol_cat2_fve, levels = c("low", "high"))
tolerance$tol_cat3_fve <- factor(tolerance$tol_cat3_fve, levels = c("low","mid", "high"))


samplelist <- read.table("output/trost/samplelist.txt", header=TRUE, sep="\t")
log_norm_ct <- read.table("output/trost/log_norm_ct.txt", header=TRUE, sep="\t")

dim(log_norm_ct)

# percentage of NAs
sum(is.na(log_norm_ct)) / (nrow(log_norm_ct)*ncol(log_norm_ct)) *100
# 0.56%
```


## Define training data
```{r define training data}
# join metadata
head(tolerance)
head(samplelist)
levels(samplelist$treatment) <- c("control", "drought stress")

# join samplelist
samplelist_joined <- join(samplelist, tolerance, by = "subspecies_id")
samplelist_pot <- droplevels(subset(samplelist_joined, samplelist_joined$cultivation == "greenhouse"))

# only pot data
log_norm_ct_pot <- subset(log_norm_ct, samplelist$cultivation=="greenhouse")

# train/test data
train_data <- subset(log_norm_ct_pot, samplelist_pot$model_set=="train")
test_data <- subset(log_norm_ct_pot, samplelist_pot$model_set=="test")

# train/test info
train_info <- droplevels(subset(samplelist_pot, samplelist_pot$model_set=="train"))
test_info <- droplevels(subset(samplelist_pot, samplelist_pot$model_set=="test"))
```


## Define control data (and meta data)
```{r define control data}
# data
train_data_control <- subset(train_data, train_info$treatment=="control")
test_data_control <- subset(test_data, test_info$treatment=="control")

# meta data
train_info_control <- droplevels(subset(train_info, train_info$treatment=="control"))
test_info_control <- droplevels(subset(test_info, test_info$treatment=="control"))
```


## Define all data (without 3 cultivars --> model_set=="NA") and subset of all control samples
```{r define all data (control)}
all_data <- subset(log_norm_ct_pot, !samplelist_pot$model_set=="NA")
all_info <- droplevels(subset (samplelist_pot, !samplelist_pot$model_set=="NA"))

# all control samples (ohne NA -> 3 Kutlivare)
all_data_control <- subset(log_norm_ct_pot, samplelist_pot$treatment=="control" & !samplelist_pot$model_set=="NA")
all_info_control <- droplevels( subset (samplelist_pot, samplelist_pot$treatment=="control" & !samplelist_pot$model_set=="NA"))
```


# Load data without NAs from PCA (rnipals, no scaling)
```{r data without NAs}
log_norm_ct_woNA <- read.table("output/trost/log_norm_ct_none_rnipals_completeObs.txt", header=TRUE, sep="\t")

# only pot data
log_norm_ct_pot_woNA <- subset(log_norm_ct_woNA, samplelist$cultivation=="greenhouse")

##### train/test #####
train_data_woNA <- subset(log_norm_ct_pot_woNA, samplelist_pot$model_set=="train")
test_data_woNA <- subset(log_norm_ct_pot_woNA, samplelist_pot$model_set=="test")

##### control ####
train_data_control_woNA <- subset(train_data_woNA, train_info$treatment=="control")
test_data_control_woNA <- subset(test_data_woNA, test_info$treatment=="control")

##### all ####
all_data_woNA <- subset(log_norm_ct_pot_woNA, !samplelist_pot$model_set=="NA")

#### all control samples #####
all_data_control_woNA <- subset(log_norm_ct_pot_woNA, samplelist_pot$treatment=="control" & !samplelist_pot$model_set=="NA")
```


# Regression tree: rpart package
## Regression tree: train data
```{r rpart reg train}
input_train <- data.frame(train_data, "tol"=train_info$mdrym_gh)

tree_train <- rpart(tol~., data=input_train)
printcp(tree_train)
plot(tree_train)
text(tree_train)

tree_train$cptable[which.min(tree_train$cptable[,"xerror"]),"CP"]
tree_train_pruned <- prune(tree_train, cp=0.1)

plot(tree_train_pruned)
text(tree_train_pruned)

# as party 
plot(as.party(tree_train_pruned))

pdf("figures/model/pot_tree_train_pruned.pdf", width=13, height=8)
plot(as.party(prune(tree_train, cp=0.16))) # 1 node
plot(as.party(prune(tree_train, cp=0.13))) # 2 nodes
plot(as.party(prune(tree_train, cp=0.12))) # 3 nodes
plot(as.party(prune(tree_train, cp=0.07))) # 4 nodes
plot(as.party(prune(tree_train, cp=0.03))) # 5 nodes
dev.off()

###############
# predict
plot(test_info$mdrym_gh, predict(tree_train, as.data.frame(test_data), type="vector"))
sqrt(mean((test_info$mdrym_gh - predict(tree_train, as.data.frame(test_data), type="vector"))^2))
```


## Regression tree: only control pot train data
```{r rpart reg train control}
input_train_control <- data.frame(train_data_control, "tol"=train_info_control$mdrym_gh)

tree_train_control <- rpart(tol~., data=input_train_control)
printcp(tree_train_control)
plot(tree_train_control)
text(tree_train_control)

tree_train_control$cptable[which.min(tree_train_control$cptable[,"xerror"]),"CP"]
tree_train_control_pruned <- prune(tree_train_control, cp=0.1639378)

plot(tree_train_control_pruned)
text(tree_train_control_pruned)

# as party 
plot(as.party(tree_train_control_pruned))

pdf("figures/model/pot_tree_train_control_pruned.pdf", width=13, height=8)
plot(as.party(prune(tree_train_control, cp=0.3))) # 1 node
plot(as.party(prune(tree_train_control, cp=0.25))) # 2 nodes
plot(as.party(prune(tree_train_control, cp=0.16))) # 3 nodes
dev.off()

###############
# predict
plot(test_info_control$mdrym_gh, predict(tree_train_control, as.data.frame(test_data_control), type="vector"))
sqrt(mean((test_info_control$mdrym_gh - predict(tree_train_control, as.data.frame(test_data_control), type="vector"))^2))
```


## Regression tree: all data
```{r rpart reg all}
input_all <- data.frame(all_data, "tol"=all_info$mdrym_gh)
sum(is.na(all_data))
# 80

tree_all <- rpart(tol~., data=input_all)
printcp(tree_all)
plot(tree_all)
text(tree_all)

tree_all$cptable[which.min(tree_all$cptable[,"xerror"]),"CP"]
tree_all_pruned <- prune(tree_all, cp=0.1634957)

plot(tree_all_pruned)
text(tree_all_pruned)

# as party 
plot(as.party(tree_all_pruned))

# pdf("figures/model/pot_tree_all_pruned.pdf", width=13, height=8)
# plot(as.party(prune(tree_all, cp=0.16))) # 1 node
# plot(as.party(prune(tree_all, cp=0.13))) # 2 nodes
# plot(as.party(prune(tree_all, cp=0.10))) # 3 nodes
# plot(as.party(prune(tree_all, cp=0.09))) # 4 nodes
# plot(as.party(prune(tree_all, cp=0.05))) # 6 nodes
# plot(as.party(prune(tree_all, cp=0.02))) # 7 nodes
# plot(as.party(prune(tree_all, cp=0.016))) # 8 nodes
# plot(as.party(prune(tree_all, cp=0.0122))) # 9 nodes
# plot(as.party(prune(tree_all, cp=0.01))) # 10 nodes
# dev.off()

###############
# predict
plot(test_info$mdrym_gh, predict(tree_all, as.data.frame(test_data), type="vector"))
sqrt(mean((test_info$mdrym_gh - predict(tree_all, as.data.frame(test_data), type="vector"))^2))

# for all data
plot(all_info$mdrym_gh, predict(tree_all, as.data.frame(all_data), type="vector"))
sqrt(mean((all_info$mdrym_gh - predict(tree_all, as.data.frame(all_data), type="vector"))^2))
```


## Regression tree: all data control
```{r rpart reg all control}
input_all_control <- data.frame(all_data_control, "tol"=all_info_control$mdrym_gh)

tree_all_control <- rpart(tol~., data=input_all_control)
printcp(tree_all_control)
plot(tree_all_control)
text(tree_all_control)

tree_all_control$cptable[which.min(tree_all_control$cptable[,"xerror"]),"CP"]
tree_all_control_pruned <- prune(tree_all_control, cp=0.09729259)

plot(tree_all_control_pruned)
text(tree_all_control_pruned)

# as party 
plot(as.party(tree_all_control_pruned))

# pdf("figures/model/pot_tree_all_control_pruned.pdf", width=13, height=8)
# plot(as.party(prune(tree_all_control, cp=0.21))) # 1 node
# plot(as.party(prune(tree_all_control, cp=0.15))) # 2 nodes
# plot(as.party(prune(tree_all_control, cp=0.10))) # 3 nodes
# plot(as.party(prune(tree_all_control, cp=0.06))) # 4 nodes
# plot(as.party(prune(tree_all_control, cp=0.05))) # 5 nodes
# plot(as.party(prune(tree_all_control, cp=0.015))) # 6 nodes
# plot(as.party(prune(tree_all_control, cp=0.01))) # 7 nodes
# dev.off()

###############
# predict
plot(test_info_control$mdrym_gh, predict(tree_all_control, as.data.frame(test_data_control), type="vector"))
sqrt(mean((test_info_control$mdrym_gh - predict(tree_all_control, as.data.frame(test_data_control), type="vector"))^2))

# for all data
plot(all_info_control$mdrym_gh, predict(tree_all_control, as.data.frame(all_data_control), type="vector"))
sqrt(mean((all_info_control$mdrym_gh - predict(tree_all_control, as.data.frame(all_data_control), type="vector"))^2))
```


# Classification tree: rpart package
## Classification tree: train data, 2 tolerance classes
```{r rpart cat2 train}
input_cat2_train <- data.frame(train_data, "tol"=train_info$tol_cat2_gh)
input_cat2_train$tol <- train_info$tol_cat2_gh

tree_cat2_train <- rpart(tol~., data=input_cat2_train)
printcp(tree_cat2_train)
plot(tree_cat2_train)
text(tree_cat2_train)

tree_cat2_train$cptable[which.min(tree_cat2_train$cptable[,"xerror"]),"CP"]
tree_cat2_train_pruned <- prune(tree_cat2_train, cp=0.175)

plot(tree_cat2_train_pruned)
text(tree_cat2_train_pruned)

# as party 
plot(as.party(tree_cat2_train_pruned))

# pdf("figures/model/pot_tree_cat2_train_pruned.pdf", width=13, height=8)
# plot(as.party(prune(tree_cat2_train, cp=0.2))) # 1 node
# plot(as.party(prune(tree_cat2_train, cp=0.09))) # 2 nodes
# plot(as.party(prune(tree_cat2_train, cp=0.05))) # 3 nodes
# plot(as.party(prune(tree_cat2_train, cp=0.02))) # 4 nodes
# dev.off()

###############
# predict
table(train_info$tol_cat2_gh, predict(tree_cat2_train, as.data.frame(train_data), type="class"))
table(test_info$tol_cat2_gh, predict(tree_cat2_train, as.data.frame(test_data), type="class"))
```


## Classification tree: train data, 3 tolerance classes
```{r rpart cat3 train}
input_cat3_train <- data.frame(train_data, "tol"=train_info$tol_cat3_gh)
input_cat3_train$tol <- train_info$tol_cat3_gh
levels(input_cat3_train$tol)

tree_cat3_train <- rpart(tol~., data=input_cat3_train)
printcp(tree_cat3_train)
plot(tree_cat3_train)
text(tree_cat3_train)

tree_cat3_train$cptable[which.min(tree_cat3_train$cptable[,"xerror"]),"CP"]
tree_cat3_train_pruned <- prune(tree_cat3_train, cp=0.01)

plot(tree_cat3_train_pruned)
text(tree_cat3_train_pruned)

# as party 
plot(as.party(tree_cat3_train_pruned))

# pdf("figures/model/pot_tree_cat3_train_pruned.pdf", width=13, height=8)
# plot(as.party(prune(tree_cat3_train, cp=0.20))) # 1 node
# plot(as.party(prune(tree_cat3_train, cp=0.10))) # 2 nodes
# plot(as.party(prune(tree_cat3_train, cp=0.08))) # 3 nodes
# plot(as.party(prune(tree_cat3_train, cp=0.05))) # 7 nodes
# dev.off()

###############
# predict
table(train_info$tol_cat3_gh, predict(tree_cat3_train, as.data.frame(train_data), type="class"))
table(test_info$tol_cat3_gh, predict(tree_cat3_train, as.data.frame(test_data), type="class"))
```

## Classification tree: all data, 3 tolerance classes
```{r rpart cat3 all}
input_cat3_all <- data.frame(all_data, "tol"=all_info$tol_cat3_gh)
input_cat3_all$tol <- all_info$tol_cat3_gh
levels(input_cat3_all$tol)

tree_cat3_all <- rpart(tol~., data=input_cat3_all)
printcp(tree_cat3_all)
plot(tree_cat3_all)
text(tree_cat3_all)

tree_cat3_all$cptable[which.min(tree_cat3_all$cptable[,"xerror"]),"CP"]
tree_cat3_all_pruned <- prune(tree_cat3_all, cp=0.01)

plot(tree_cat3_all_pruned)
text(tree_cat3_all_pruned)

# as party 
plot(as.party(tree_cat3_all_pruned))

# pdf("figures/model/pot_tree_cat3_all_pruned.pdf", width=13, height=8)
# plot(as.party(prune(tree_cat3_all, cp=0.20))) # 1 node
# plot(as.party(prune(tree_cat3_all, cp=0.13))) # 2 nodes
# plot(as.party(prune(tree_cat3_all, cp=0.08))) # 3 nodes
# plot(as.party(prune(tree_cat3_all, cp=0.05))) # 7 nodes
# dev.off()

###############
# predict
table(all_info$tol_cat3_gh, predict(tree_cat3_all, as.data.frame(all_data), type="class"))
table(test_info$tol_cat3_gh, predict(tree_cat3_all, as.data.frame(test_data), type="class"))
```


# Random forest regression
## Random forest regression: train data
```{r rf reg train}
input_train_woNA <- data.frame(train_data_woNA, "tol"=train_info$mdrym_gh)

rf_train <- randomForest(tol~. , data=input_train_woNA, ntree=1000)
print(rf_train)

#importance(rf_train)
varImpPlot(rf_train)

plot(train_info$mdrym_gh, rf_train$predicted)
abline(0,1)

plot(test_info$mdrym_gh, predict(rf_train, test_data_woNA))
sqrt(mean((test_info$mdrym_gh - predict(rf_train, as.data.frame(test_data_woNA)))^2))
```


### Cross-Validation
```{r rf reg train CV}
set.seed(3)
rf_train_cv <- rfcv(train_data_woNA, train_info$mdrym_gh, step=0.8)
rf_train_cv$n.var
with(rf_train_cv, plot(n.var, error.cv, log="x", type="o", lwd=2))
```


## Random forest regression: train control data
```{r rf reg train control}
input_train_control_woNA <- data.frame(train_data_control_woNA, "tol"=train_info_control$mdrym_gh)

rf_train_control <- randomForest(tol~. , data=input_train_control_woNA, ntree=1000)
print(rf_train_control)

#importance(rf_train_control)
varImpPlot(rf_train_control)

plot(train_info_control$mdrym_gh, rf_train_control$predicted)
abline(0,1)

plot(test_info_control$mdrym_gh, predict(rf_train_control, test_data_control_woNA))
sqrt(mean((test_info_control$mdrym_gh - predict(rf_train_control, as.data.frame(test_data_control_woNA)))^2))
```


## Random forest regression: train data
```{r rf reg all}
input_all_woNA <- data.frame(all_data_woNA, "tol"=all_info$mdrym_gh)

rf_all <- randomForest(tol~. , data=input_all_woNA, ntree=1000)
print(rf_all)

#importance(rf_all)
varImpPlot(rf_all)

plot(all_info$mdrym_gh, rf_all$predicted)
abline(0,1)

plot(test_info$mdrym_gh, predict(rf_all, test_data_woNA))
sqrt(mean((test_info$mdrym_gh - predict(rf_all, as.data.frame(test_data_woNA)))^2))
```

### Cross-Validation
```{r rf reg all CV}
set.seed(3)
rf_all_cv <- rfcv(all_data_woNA, all_info$mdrym_gh, step=0.8)
rf_all_cv$n.var
with(rf_all_cv, plot(n.var, error.cv, log="x", type="o", lwd=2))
```


## Random forest regression: train control data
```{r rf reg all control}
input_all_control_woNA <- data.frame(all_data_control_woNA, "tol"=all_info_control$mdrym_gh)

rf_all_control <- randomForest(tol~. , data=input_all_control_woNA, ntree=1000)
print(rf_all_control)

#importance(rf_all_control)
varImpPlot(rf_all_control)

plot(all_info_control$mdrym_gh, rf_all_control$predicted)
abline(0,1)

plot(test_info_control$mdrym_gh, predict(rf_all_control, test_data_control_woNA))
sqrt(mean((test_info_control$mdrym_gh - predict(rf_all_control, as.data.frame(test_data_control_woNA)))^2))
```


# Random forest classification
## Random forest classification: train data with 2 classes
```{r rf cat2 train}
input_cat2_train_woNA <- data.frame(train_data_woNA, "tol"=train_info$tol_cat2_gh)
input_cat2_train_woNA$tol <- train_info$tol_cat2_gh
levels(input_cat2_train_woNA$tol)

rf_cat2_train <- randomForest(tol~. , data=input_cat2_train_woNA, ntree=1000)
print(rf_cat2_train)

write.table(importance(rf_cat2_train), "output/model/importance_pot_rf_cat2_train.txt", sep="\t")
varImpPlot(rf_cat2_train)

table(predict(rf_cat2_train, test_data_woNA), test_info$tol_cat2_gh)
confusionMatrix(table(predict(rf_cat2_train, test_data_woNA), test_info$tol_cat2_gh))
```


## Random forest classification: train data with 3 classes
```{r rf cat3 train}
input_cat3_train_woNA <- data.frame(train_data_woNA, "tol"=train_info$tol_cat3_gh)
input_cat3_train_woNA$tol <- train_info$tol_cat3_gh
levels(input_cat3_train_woNA$tol)

rf_cat3_train <- randomForest(tol~. , data=input_cat3_train_woNA, ntree=1000)
print(rf_cat3_train)

write.table(importance(rf_cat3_train), "output/model/importance_pot_rf_cat3_train.txt", sep="\t")
varImpPlot(rf_cat3_train)

table(predict(rf_cat3_train, test_data_woNA), test_info$tol_cat3_gh)
confusionMatrix(table(predict(rf_cat3_train, test_data_woNA), test_info$tol_cat3_gh))
```


### Cross-Validation
```{r rf cat3 train CV}
set.seed(1)
rf_cat3_train_cv <- rfcv(train_data_woNA, train_info$tol_cat3_gh, step=0.8)
rf_cat3_train_cv$n.var
with(rf_cat3_train_cv, plot(n.var, error.cv, log="x", type="o", lwd=2))
```


### Variable Selection (vs)
```{r rf cat3 train VarSel}
set.seed(1)
rf_cat3_train_vs <- varSelRF(train_data_woNA, train_info$tol_cat3_gh, ntree = 500, ntreeIterat = 300, vars.drop.frac = 0.2, c.sd=1)
rf_cat3_train_vs
# 7
write.table(rf_cat3_train_vs$selected.vars, "output/model/pot_rf_cat3_train_selected_vars.txt", sep="\t")
rf_cat3_train_vs$selected.vars
plot(rf_cat3_train_vs, which=1)
plot(rf_cat3_train_vs, which=2)

# indices of selected variables
vs_idx <- which(colnames(train_data_woNA) %in% rf_cat3_train_vs$selected.vars)
```


## Random forest classification: selected variables from pot data with 3 classes
```{r rf cat3 train selected variables, fig.width=12}
train_data_woNA_vs <- train_data_woNA[,vs_idx]
test_data_woNA_vs <- test_data_woNA[,vs_idx]

input_train_cat3_woNA_vs <- data.frame(train_data_woNA_vs, "tol"=train_info$tol_cat3_gh)
input_train_cat3_woNA_vs$tol <- train_info$tol_cat3_gh
levels(input_train_cat3_woNA_vs$tol)

rf_cat3_train_reduced <- randomForest(tol~. , data=input_train_cat3_woNA_vs, ntree=1000)
print(rf_cat3_train_reduced)

write.table(importance(rf_cat3_train_reduced), "output/model/importance_rf_cat3_train_reduced.txt", sep="\t")
par(mfrow=c(1,2))
varImpPlot(rf_cat3_train_reduced)
varImpPlot(rf_cat3_train)
par(mfrow=c(1,1))

table(predict(rf_cat3_train_reduced, test_data_woNA_vs), test_info$tol_cat3_gh)
confusionMatrix(table(predict(rf_cat3_train_reduced, test_data_woNA_vs), test_info$tol_cat3_gh))
# full model
confusionMatrix(table(predict(rf_cat3_train, test_data_woNA), test_info$tol_cat3_gh))
```


## Random forest classification: train control data with 3 classes
```{r rf cat3 train control}
input_cat3_train_control_woNA <- data.frame(train_data_control_woNA, "tol"=train_info_control$tol_cat3_gh)
input_cat3_train_control_woNA$tol <- train_info_control$tol_cat3_gh
levels(input_cat3_train_control_woNA$tol)

rf_cat3_train_control <- randomForest(tol~. , data=input_cat3_train_control_woNA, ntree=1000)
print(rf_cat3_train_control)

write.table(importance(rf_cat3_train_control), "output/model/importance_pot_rf_cat3_train_control.txt", sep="\t")
varImpPlot(rf_cat3_train_control)

table(predict(rf_cat3_train_control, test_data_control_woNA), test_info_control$tol_cat3_gh)
confusionMatrix(table(predict(rf_cat3_train_control, test_data_control_woNA), test_info_control$tol_cat3_gh))
```


## Random forest classification: all data with 3 classes
```{r rf cat3 all}
input_cat3_all_woNA <- data.frame(all_data_woNA, "tol"=all_info$tol_cat3_gh)
input_cat3_all_woNA$tol <- all_info$tol_cat3_gh
levels(input_cat3_all_woNA$tol)

rf_cat3_all <- randomForest(tol~. , data=input_cat3_all_woNA, ntree=1000)
print(rf_cat3_all)

write.table(importance(rf_cat3_all), "output/model/importance_pot_rf_cat3_all.txt", sep="\t")
varImpPlot(rf_cat3_all)

table(all_info$tol_cat3_gh, rf_cat3_all$predicted)
```


### Cross-Validation
```{r rf cat3 all CV}
set.seed(5)
rf_cat3_all_cv <- rfcv(all_data_woNA, all_info$tol_cat3_gh, step=0.8)
rf_cat3_all_cv$n.var
with(rf_cat3_all_cv, plot(n.var, error.cv, log="x", type="o", lwd=2))
```


### Variable Selection (vs)
```{r rf cat3 all VarSel}
set.seed(1)
rf_cat3_all_vs <- varSelRF(all_data_woNA, all_info$tol_cat3_gh, ntree = 500, ntreeIterat = 300, vars.drop.frac = 0.2, c.sd=1)
rf_cat3_all_vs
# 18
write.table(rf_cat3_all_vs$selected.vars, "output/model/pot_rf_cat3_all_selected_vars.txt", sep="\t")
rf_cat3_all_vs$selected.vars
plot(rf_cat3_all_vs, which=1)
plot(rf_cat3_all_vs, which=2)

```


## Random forest classification: all control data with 3 classes
```{r rf cat3 all control}
input_cat3_all_control_woNA <- data.frame(all_data_control_woNA, "tol"=all_info_control$tol_cat3_gh)
input_cat3_all_control_woNA$tol <- all_info_control$tol_cat3_gh
levels(input_cat3_all_control_woNA$tol)

rf_cat3_all_control <- randomForest(tol~. , data=input_cat3_all_control_woNA, ntree=1000)
print(rf_cat3_all_control)

write.table(importance(rf_cat3_all_control), "output/model/importance_pot_rf_cat3_all_control.txt", sep="\t")
varImpPlot(rf_cat3_all_control)

table(all_info_control$tol_cat3_gh, rf_cat3_all_control$predicted)
```


# Export importance tables
```{r export importance tables}
importance_rf_reg <- data.frame(importance(rf_train), 
                            importance(rf_train_control), 
                            importance(rf_all), 
                            importance(rf_all_control))
colnames(importance_rf_reg) <- c("reg_train", "reg_train_control", "reg_all", "reg_all_control")
head(importance_rf_reg)

importance_rf_cat3 <- data.frame(importance(rf_cat3_train), 
                            importance(rf_cat3_train_control), 
                            importance(rf_cat3_all), 
                            importance(rf_cat3_all_control))
colnames(importance_rf_cat3) <- c("cat3_train", "cat3_train_control", "cat3_all", "cat3_all_control")
head(importance_rf_cat3)

plot(importance_rf_cat3$cat3_train, importance_rf_cat3$cat3_train_control)
plot(importance_rf_cat3$cat3_train, importance_rf_cat3$cat3_all)

write.table(importance_rf_reg, "output/model/pot_importance_rf_reg.txt", sep="\t")
write.table(importance_rf_cat3, "output/model/pot_importance_rf_cat3.txt", sep="\t")
```


# Predict field DRYM
## Random forest classification: train data with 3 classes
```{r rf cat3 train FIELD TOL}
input_cat3_train_woNA_field <- data.frame(train_data_woNA, "tol"=train_info$tol_cat3_fve)
input_cat3_train_woNA_field$tol <- train_info$tol_cat3_fve
levels(input_cat3_train_woNA_field$tol)

set.seed(1234)
rf_cat3_train_field <- randomForest(tol~. , data=input_cat3_train_woNA_field, ntree=1000)
print(rf_cat3_train_field)

table(predict(rf_cat3_train_field, test_data_woNA), test_info$tol_cat3_fve)
confusionMatrix(table(predict(rf_cat3_train_field, test_data_woNA), test_info$tol_cat3_fve))

write.table(importance(rf_cat3_train_field), "output/model/importance_greenhouse_rf_cat3_train_field_tol.txt", sep="\t")
```


### Cross-Validation
```{r rf cat3 train CV FIELD TOL}
set.seed(1)
rf_cat3_train_cv_field <- rfcv(train_data_woNA, train_info$tol_cat3_fve, step=0.8)
length(rf_cat3_train_cv_field$n.var)
with(rf_cat3_train_cv_field, plot(n.var, error.cv, log="x", type="o", lwd=2))

#pdf("../../../../Doktorarbeit/figures/model_greenhouse_transcripts_rfcv_cat3_train_field.pdf", width=6, height=6)
pdf("figures/thesis/model_greenhouse_transcripts_rfcv_cat3_train_field.pdf", width=6, height=6)
par(mar=c(4.5, 4.5, 0.5, 0.6))
with(rf_cat3_train_cv_field, plot(n.var, error.cv, log="x", type="o", lwd=2, cex.lab=1.5, cex.axis=1.2,
                            xlab="number of predictors", ylab="5-fold cross-validation error"))
dev.off()
```


### Variable Selection (vs)
```{r rf cat3 train VarSel FIELD TOL}
set.seed(1)
rf_cat3_train_vs_field <- varSelRF(train_data_woNA, train_info$tol_cat3_fve, ntree = 500, ntreeIterat = 300, vars.drop.frac = 0.2, c.sd=1)
rf_cat3_train_vs_field
# 18
#write.table(rf_cat3_train_vs$selected.vars, "output/model/pot_rf_cat3_train_selected_vars.txt", sep="\t")
rf_cat3_train_vs_field$selected.vars
plot(rf_cat3_train_vs_field, which=1)
plot(rf_cat3_train_vs_field, which=2)

# indices of selected variables
vs_idx_field <- which(colnames(train_data_woNA) %in% rf_cat3_train_vs_field$selected.vars)
```


## Random forest classification: selected variables from pot data with 3 classes
```{r rf cat3 train selected variables FIELD TOL}
train_data_woNA_vs_field <- train_data_woNA[,vs_idx_field]
test_data_woNA_vs_field <- test_data_woNA[,vs_idx_field]

input_train_cat3_woNA_vs_field <- data.frame(train_data_woNA_vs_field, "tol"=train_info$tol_cat3_fve)
input_train_cat3_woNA_vs_field$tol <- train_info$tol_cat3_fve
levels(input_train_cat3_woNA_vs_field$tol)

set.seed(1)
rf_cat3_train_reduced_field <- randomForest(tol~. , data=input_train_cat3_woNA_vs_field, ntree=1000)
print(rf_cat3_train_reduced_field)

#write.table(importance(rf_cat3_train_reduced_field), "output/model/importance_rf_cat3_train_reduced_field.txt", sep="\t")

table(predict(rf_cat3_train_reduced_field, test_data_woNA_vs_field), test_info$tol_cat3_fve)
confusionMatrix(table(predict(rf_cat3_train_reduced_field, test_data_woNA_vs_field), test_info$tol_cat3_fve))
# full model
confusionMatrix(table(predict(rf_cat3_train_field, test_data_woNA), test_info$tol_cat3_fve))
```


# PLS
## Step 1. PLS (Partial Least Squares) for all variables to compute VIP
```{r pls all}
input_all_woNA <- data.frame(train_data_woNA, "tol"=train_info$mdrym_gh)

pls_all <- plsr(tol ~ ., data=input_all_woNA, validation="LOO", method="oscorespls")

summary(pls_all)

par(mfrow=c(1,2))
plot(pls_all, "validation", estimate="CV")
par(pty="s")
plot(pls_all, "prediction", ncomp=4)
abline(0,1, col="gray")
par(mfrow=c(1,1))

plot(RMSEP(pls_all), legendpos="topright") # Plot the root mean squared error of prediction (RMSEP) to get the number of components to be used

#RMSEP(pls_all)
M.pls <- which.min(RMSEP(pls_all, estimate="adjCV")$val) - 1
# 12 --> 11 comps!
plot(pls_all, "validation", estimate="adjCV")
abline(v=11, col="red")

which.max(R2(pls_all)$val)
plot(R2(pls_all))
# 12 --> 11 comps!
abline(v=11, col="red")

test_data_woNA_2 <- data.frame(test_data_woNA, "tol"=test_info$mdrym_gh)
RMSEP(pls_all, ncomp=1, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=2, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=3, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=4, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=5, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=6, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=7, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=8, newdata=test_data_woNA_2, intercept=FALSE)
RMSEP(pls_all, ncomp=9, newdata=test_data_woNA_2, intercept=FALSE)

# The interpretation of the scores and loadings is similar to PCA: a score indicates how much a particular object contributes to a latent variable, while a loading indicates the contribution of a particular variable. An example of a loading plot is obtained using the code below:
plot(pls_all, "loading", comps = 1:3, legendpos = "top", lty = c(1, 2, 4), col = c(1, 2, 4))

y.est=predict(pls_all, newdata=test_data_woNA_2, ncomp=M.pls)
cor(y.est, test_info$mdrym_gh)
```

## VIP function
```{r VIP function}
VIPjh=function(object, j, h)
{
   if (object$method!="oscorespls")
      stop("Only implemented for orthogonal scores algorithm. Refit with 'method = \"oscorespls\"'")
   if (nrow(object$Yloadings) > 1)
      stop("Only implemented for single-response models")

   b=c(object$Yloadings)[1:h]
   T=object$scores[,1:h, drop=F]
   SS=b^2 * colSums(T^2)
   W=object$loading.weights[,1:h, drop=F]
   Wnorm2=colSums(W^2)
   sqrt(nrow(W) * sum(SS * W[j,]^2 / Wnorm2) / sum(SS))
}
```


## Step 2. Compute VIP-value for all variables and Step 3. Sort them accordingly
```{r compute VIP values}
p <- ncol(train_data)
VIP=vector("numeric", p)
I.pls=1:p

for(j in 1:p)
   VIP[j]=VIPjh(pls_all, j, M.pls)

# Plot result
#pdf("Step_2.pdf")
plot(VIP)
#dev.off()

VIP_table <- cbind(colnames(train_data), VIP)
head(VIP_table[order(VIP_table[,2], decreasing=T),])

VIPmax=sort(VIP, decreasing=T)
for (i in 1:p)
I.pls[i]=which(VIP==VIPmax[i])
```


# BioMark package
```{r BioMark}
# stability_all <- get.biom(X = train_data_woNA, Y = train_info$mdrym_gh, fmethod=c("pcr", "pls", "vip", "lasso"), type = "stab")
# summary(stability_all)
# selection(stability_all)
# 
# # only two class discrimination possible!
# HC_cat2 <- get.biom(X = train_data_woNA, Y = train_info$tol_cat2_gh, fmethod=c("studentt","pcr", "pls", "vip"), type = "HC")
# summary(HC_cat2)
# selection(HC_cat2)
# 
# selection(HC_cat2)$pls[[1]]
# selection(HC_cat2)$vip[[1]]
# selection(HC_cat2)$pcr[[1]]
# intersect( selection(HC_cat2)$pls[[1]], selection(HC_cat2)$vip[[1]])
# intersect( selection(HC_cat2)$pls[[1]], selection(HC_cat2)$pcr[[1]])
# colnames(train_data_control_woNA)
```


# Save workspace and sessionInfo
```{r save workspace}
save.image("qpcr_model_pot.RData")
sessionInfo()
```

