---
title: "Prediction Model using qPCR data (TROST field) to validate breeder trials"
author: "Heike Sprenger"
date: "Wednesday, March 30, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
---


# Set working directory  
```{r set working directory}
getwd()
#setwd("D:/work/repos/trost_qpcr/")
#setwd("~/work/repos/trost_qpcr/")
```


# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(reshape)
library(pander)
library(pcaMethods)
library(ggplot2)
library(corrplot)
library(gplots)
library(RColorBrewer)
library(plyr)
library(randomForest)
library(caret) 
library(varSelRF)
library(glmnet)
library(mixOmics)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("qpcr_model_field_breeder.RData")
```


# Load tolerance data
```{r load tolerance data}
tolerance <- read.table("../trost_phenotypes/output/tolerance_phenotyper_metadata_2sub.txt", 
                        sep = "\t", header = TRUE)

# relevel tolerance factors
tolerance$tol_cat2_fve <- factor(tolerance$tol_cat2_fve, levels = c("low", "high"))
tolerance$tol_cat3_fve <- factor(tolerance$tol_cat3_fve, levels = c("low","mid", "high"))
```


# Load qPCR samplelist
* ``log_norm_ct_all`` contains data from 423 samples (TROST Field + Breeder) and 43 genes, **with NAs!**
* comes from: ``qpcr_data_field_and_breeder.Rmd``
* ``samplelist_all`` is ordered by sampleID

```{r Load qPCR samplelist}
# TROST + Breeder samplelist
samplelist_all <- read.table("output/model/samplelist_field_breeder.txt", header = T, sep = "\t")
dim(samplelist_all)
head(samplelist_all)

# join samplelist with tolerance information
samplelist_joined <- join(samplelist_all, tolerance, by = "subspecies_id")
head(samplelist_joined)
dim(samplelist_joined)

# data contains NAs!
# log_norm_ct_all <- read.table("output/model/log_norm_ct_field_breeder.txt", header = T, sep = "\t")
```


## Define subset for Field and Breeder experiments
```{r define subset for field and breeder experiments}
# samplelist only for TROST FIELD data
samplelist_field_joined <- droplevels(subset(samplelist_joined, 
                                             samplelist_joined$trost_breeder == "field"))
dim(samplelist_field_joined)


# samplelist only for BREEDER data
samplelist_breeder_joined <- droplevels(subset(samplelist_joined, 
                                               samplelist_joined$trost_breeder == "breeder"))
dim(samplelist_breeder_joined)
```


## Define FIELD train/test info
```{r define FIELD train/test info}
# train/test info
train_info <- subset(samplelist_field_joined, samplelist_field_joined$model_set=="train")
test_info <- subset(samplelist_field_joined, samplelist_field_joined$model_set=="test")

# only control samples
train_info_control <- droplevels(subset(train_info, train_info$treatment == "control"))
test_info_control <- droplevels(subset(test_info, test_info$treatment == "control"))
```


## Define all FIELD sample information (without 3 cultivars --> model_set=="NA") and subset of all FIELD control samples
```{r define all FIELD sample information (control)}
# all TROST samples (without NA --> 3 cultivars)
all_info <- subset (samplelist_field_joined, !samplelist_field_joined$model_set=="NA")
dim(all_info)

# all TROST control samples (without NA --> 3 cultivars)
all_info_control <- subset (samplelist_field_joined, 
                            samplelist_field_joined$treatment=="control" & !samplelist_field_joined$model_set=="NA")
```


## Define all Breeder sample information (without 3 cultivars --> model_set=="NA")
```{r define all breeder sample information}
# all BREEDER samples without NA --> 3 cultivars
all_breeder_info <- droplevels(subset (samplelist_breeder_joined, !samplelist_breeder_joined$model_set=="NA"))
dim(all_breeder_info)

# years 2011/2012
all_breeder_info_2011 <- droplevels( subset(all_breeder_info, all_breeder_info$treatment == "2011"))
all_breeder_info_2012 <- droplevels( subset(all_breeder_info, all_breeder_info$treatment == "2012"))

# information for separate trials or combinations
get_subset_breeder <- function(x = all_breeder_info, trial_name) {
  return( droplevels( subset(x, x$trial == trial_name) ) ) }

all_info_norika_2011 <- get_subset_breeder(trial_name = "Norika_2011")
all_info_norika_2012 <- get_subset_breeder(trial_name = "Norika_2012")
all_info_petersgroden_2011 <- get_subset_breeder(trial_name = "Petersgroden_2011")
all_info_petersgroden_2012 <- get_subset_breeder(trial_name = "Petersgroden_2012")
all_info_windeby_2011 <- get_subset_breeder(trial_name = "Windeby_2011")
all_info_windeby_2012 <- get_subset_breeder(trial_name = "Windeby_2012")

# locations
get_subset_location <- function(x = all_breeder_info, loc_name) {
  return( droplevels( subset(x, x$cultivation == loc_name) ) ) }

all_info_norika <- get_subset_location(loc_name = "Norika")
all_info_petersgroden <- get_subset_location(loc_name = "Petersgroden")
all_info_windeby <- get_subset_location(loc_name = "Windeby")
```


# Load data without NAs from PCA (rnipals, no scaling, completeObs)
* comes from: ``qpcr_data_field_and_breeder.Rmd``
* ``compObs``: rownames contain ordered sampleIDs, colnames are DMT IDs
* field subset: ``compObs_field``

```{r load data without NAs}
# un-normalized data
compObs <- read.table("output/model/log_norm_ct_prep_none_rnipals_completeObs_field_breeder.txt",
                                   header = T, sep = "\t")

# ANOVA-normalized data
# compObs <- read.table("output/model/log_norm_ct_2_prep_none_rnipals_completeObs_field_breeder.txt",
#                                   header = T, sep = "\t")

dim(compObs)
# 387 43

#### only TROST FIELD data ####
compObs_field <- subset(compObs, samplelist_all$trost_breeder == "field")
dim(compObs_field)
# 202 43

#### only Breeder data #####
compObs_breeder <- subset(compObs, samplelist_all$trost_breeder == "breeder")
dim(compObs_breeder)
# 185 43
```


## Define train and test data (for field experiments)
```{r define train/test data}
##### train/test #####
train_data <- subset(compObs_field, samplelist_field_joined$model_set=="train")
test_data <- subset(compObs_field, samplelist_field_joined$model_set=="test")
dim(train_data)
dim(test_data)

##### control ####
train_data_control <- subset(train_data, train_info$treatment=="control")
test_data_control <- subset(test_data, test_info$treatment=="control")
```


## Define ALL data (without 3 cultivars --> model_set=="NA") and subset of all control samples
```{r define ALL data}
##### all (TROST field) ####
all_data <- subset(compObs_field, !samplelist_field_joined$model_set=="NA")
dim(all_data)
# 202 43

#### all control samples #####
all_data_control <- subset(compObs_field, 
                           samplelist_field_joined$treatment == "control" & !samplelist_field_joined$model_set == "NA")
```


## Define BREEDER data
* breeder subset: ``compObs_breeder``

```{r define BREEDER data}
# without 3 cultivars --> model_set=="NA"
all_breeder_data <- subset(compObs_breeder, !samplelist_breeder_joined$model_set=="NA")
dim(all_breeder_data)
# 185 43

# years 2011/2012
all_breeder_data_2011 <- subset(all_breeder_data, all_breeder_info$treatment == "2011")
all_breeder_data_2012 <- subset(all_breeder_data, all_breeder_info$treatment == "2012")

# information for separate trials or combinations
get_subset_breeder2 <- function(x = all_breeder_data, y = all_breeder_info, trial_name) {
  return( droplevels( subset(x, y$trial == trial_name) ) ) }

all_data_norika_2011 <- get_subset_breeder2(trial_name = "Norika_2011")
all_data_norika_2012 <- get_subset_breeder2(trial_name = "Norika_2012")
all_data_petersgroden_2011 <- get_subset_breeder2(trial_name = "Petersgroden_2011")
all_data_petersgroden_2012 <- get_subset_breeder2(trial_name = "Petersgroden_2012")
all_data_windeby_2011 <- get_subset_breeder2(trial_name = "Windeby_2011")
all_data_windeby_2012 <- get_subset_breeder2(trial_name = "Windeby_2012")

# locations
all_data_norika <- subset(all_breeder_data, all_breeder_info$cultivation == "Norika")
all_data_petersgroden <- subset(all_breeder_data, all_breeder_info$cultivation == "Petersgroden")
all_data_windeby <- subset(all_breeder_data, all_breeder_info$cultivation == "Windeby")
```


# Random forest regression
## Random forest regression: train data
```{r rf reg train data}
input_train <- data.frame(train_data, "tol" = train_info$mdrym_fve, check.names = FALSE)
class(input_train$tol)

i <- match("tol", names(input_train)) # i is index of tol column

set.seed(1)
rf_train <- randomForest(input_train[,-i], input_train[,i], ntree=1000)
print(rf_train)
cor.test(rf_train$predicted, train_info$mdrym_fve)

varImpPlot(rf_train)
tail(importance(rf_train)[order(importance(rf_train)),])


plot(test_info$mdrym_fve, predict(rf_train, test_data))
cor.test(test_info$mdrym_fve, predict(rf_train, test_data))
sqrt(mean((test_info$mdrym_fve - predict(rf_train, test_data))^2))
```


## Random forest regression: train control data
```{r rf reg train control data}
input_train_control <- data.frame(train_data_control, 
                                  "tol" = train_info_control$mdrym_fve, check.names = FALSE)
class(input_train_control$tol)

i <- match("tol", names(input_train_control)) # i is index of tol column

set.seed(1)
rf_train_control <- randomForest(input_train_control[,-i], input_train_control[,i], ntree=1000)
print(rf_train_control)
cor.test(rf_train_control$predicted, train_info_control$mdrym_fve)

varImpPlot(rf_train_control)

plot(test_info_control$mdrym_fve, predict(rf_train_control, test_data_control))
cor.test(test_info_control$mdrym_fve, predict(rf_train_control, test_data_control))
sqrt(mean((test_info_control$mdrym_fve - predict(rf_train_control, test_data_control))^2))
```


## Random forest regression: all data
```{r rf reg all}
input_all <- data.frame(all_data, "tol" = all_info$mdrym_fve)

set.seed(1)
rf_all <- randomForest(tol~. , data = input_all, ntree = 1000)
print(rf_all)

#importance(rf_all)
varImpPlot(rf_all)

plot(all_info$mdrym_fve, rf_all$predicted)
abline(0,1)
cor.test(all_info$mdrym_fve, rf_all$predicted)
cor.test(all_info$mdrym_fve, predict(rf_all, all_data))

plot(test_info$mdrym_fve, predict(rf_all, test_data))
sqrt(mean((test_info$mdrym_fve - predict(rf_all, as.data.frame(test_data)))^2))
```


## Random forest regression: all control data
```{r rf reg all control}
input_all_control <- data.frame(all_data_control, "tol" = all_info_control$mdrym_fve)

set.seed(1)
rf_all_control <- randomForest(tol~. , data = input_all_control, ntree = 1000)
print(rf_all_control)

varImpPlot(rf_all_control)

plot(all_info_control$mdrym_fve, rf_all_control$predicted)
abline(0,1)
cor.test(all_info_control$mdrym_fve, rf_all_control$predicted)
cor.test(all_info_control$mdrym_fve, predict(rf_all_control, all_data_control))
```


## Random forest regression: predict DRYM for breeder data (Overview)
```{r rf reg predict DRYM for breeder data OVERVIEW}
# ALL data model
cor.test(all_breeder_info$mdrym_fve, predict(rf_all, all_breeder_data))
# cor: 0.5276

# ALL CONTROL model
cor.test(all_breeder_info$mdrym_fve, predict(rf_all_control, all_breeder_data))
# cor: 0.4942314

# TRAIN data model
cor.test(all_breeder_info$mdrym_fve, predict(rf_train, all_breeder_data))
# cor: 0.3908902
plot(all_breeder_info$mdrym_fve, predict(rf_train, all_breeder_data))

# TRAIN control data model
cor.test(all_breeder_info$mdrym_fve, predict(rf_train_control, all_breeder_data))
# cor: 0.3722429
```


## Random forest regression: Predict DRYM for single breeder trials
```{r rf reg all predict DRYM for single breeder trials}
# ALL data model in detail
rf_all_pred_breeder <- predict(rf_all, all_breeder_data)
rf_all_pred_breeder_2011 <- predict(rf_all, all_breeder_data_2011)
rf_all_pred_breeder_2012 <- predict(rf_all, all_breeder_data_2012)

plot(all_breeder_info$mdrym_fve, rf_all_pred_breeder, 
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19,
     xlim = c(-0.06, 0.1), ylim = c(-0.06, 0.1))
abline(0, 1, col = "red")

cor.test(all_breeder_info$mdrym_fve, rf_all_pred_breeder)
sqrt(mean((all_breeder_info$mdrym_fve - rf_all_pred_breeder)^2))

# 2011/12
cor.test(all_breeder_info_2011$mdrym_fve, rf_all_pred_breeder_2011)$estimate # 0.45
cor.test(all_breeder_info_2012$mdrym_fve, rf_all_pred_breeder_2012)$estimate # 0.68

# Norika 2011/2012
rf_all_pred_norika_2011 <- predict(rf_all, all_data_norika_2011)
rf_all_pred_norika_2012 <- predict(rf_all, all_data_norika_2012)

plot(all_info_norika_2011$mdrym_fve, rf_all_pred_norika_2011, xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19)
text(all_info_norika_2011$mdrym_fve, rf_all_pred_norika_2011, labels = all_info_norika_2011$cultivar, cex = 0.6)
plot(all_info_norika_2012$mdrym_fve, rf_all_pred_norika_2012, xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19)
text(all_info_norika_2012$mdrym_fve, rf_all_pred_norika_2012, labels = all_info_norika_2012$cultivar, cex = 0.6)

cor(all_info_norika_2011$mdrym_fve, predict(rf_all, all_data_norika_2011))
cor(all_info_norika_2012$mdrym_fve, predict(rf_all, all_data_norika_2012))
cor(all_info_petersgroden_2011$mdrym_fve, predict(rf_all, all_data_petersgroden_2011))
cor(all_info_petersgroden_2012$mdrym_fve, predict(rf_all, all_data_petersgroden_2012))
cor(all_info_windeby_2011$mdrym_fve, predict(rf_all, all_data_windeby_2011))
cor(all_info_windeby_2012$mdrym_fve, predict(rf_all, all_data_windeby_2012))
```


# Random forest classification
## Random forest classification: train data with 3 classes
```{r rf cat3 train}
input_cat3_train <- data.frame(train_data, "tol"=train_info$tol_cat3_fve)
levels(input_cat3_train$tol)

set.seed(1)
rf_cat3_train <- randomForest(tol~. , data=input_cat3_train, ntree=1000)
print(rf_cat3_train)

varImpPlot(rf_cat3_train)
```


### Cross-Validation
```{r rf cat3 train CV}
set.seed(1)
rf_cat3_train_cv <- rfcv(train_data, train_info$tol_cat3_fve, step=0.8)
rf_cat3_train_cv$n.var
length(rf_cat3_train_cv$n.var)
#  15
with(rf_cat3_train_cv, plot(n.var, error.cv, log="x", type="o", lwd=2))
```


## Random forest classification: train_control data with 3 classes
```{r rf cat3 train control}
# only CONTROL data
input_cat3_train_control <- data.frame(train_data_control, "tol"=train_info_control$tol_cat3_fve)
levels(input_cat3_train_control$tol)

set.seed(1)
rf_cat3_train_control <- randomForest(tol~. , data=input_cat3_train_control, ntree=1000)
print(rf_cat3_train_control)
```


## Random forest classification: all data with 3 classes
```{r rf cat3 all}
input_cat3_all <- data.frame(all_data, "tol"=all_info$tol_cat3_fve)
levels(input_cat3_all$tol)

set.seed(1)
rf_cat3_all <- randomForest(tol~. , data=input_cat3_all, ntree=1000)
print(rf_cat3_all)

#write.table(importance(rf_cat3_all), "output/model/importance_field_rf_cat3_all.txt", sep="\t")
varImpPlot(rf_cat3_all)
#table(all_info$tol_cat3_fve, rf_cat3_all$predicted)
```


### Cross-Validation
```{r rf cat3 all CV}
set.seed(13)
rf_cat3_all_cv <- rfcv(all_data, all_info$tol_cat3_fve, step=0.8)
rf_cat3_all_cv$n.var
length(rf_cat3_all_cv$n.var)
# 15
with(rf_cat3_all_cv, plot(n.var, error.cv, log="x", type="o", lwd=2))
```


### Variable Selection (vs)
```{r rf cat3 all VarSel}
set.seed(1)
rf_cat3_all_vs <- varSelRF(all_data, all_info$tol_cat3_fve, 
                           ntree = 500, ntreeIterat = 300, 
                           vars.drop.frac = 0.2, c.sd=1)
rf_cat3_all_vs
# 14 transcripts

#write.table(rf_cat3_all_vs$selected.vars, "output/model/field_rf_cat3_all_selected_vars.txt", sep="\t")
#rf_cat3_all_vs$selected.vars
plot(rf_cat3_all_vs, which=1)
plot(rf_cat3_all_vs, which=2)

# indices of selected variables
rf_cat3_all_vs_idx <- which(colnames(all_data) %in% rf_cat3_all_vs$selected.vars)
```


## Random forest classification: selected variables from all data with 3 classes
```{r rf cat3 all selected variables, fig.width=12}
all_data_vs <- all_data[,rf_cat3_all_vs_idx]
dim(all_data_vs)
# 202 14
all_breeder_data_vs <- all_breeder_data[,rf_cat3_all_vs_idx]

input_all_cat3_vs <- data.frame(all_data_vs, "tol"=all_info$tol_cat3_fve)
levels(input_all_cat3_vs$tol)

set.seed(1)
rf_cat3_all_reduced <- randomForest(tol~. , data=input_all_cat3_vs, ntree=1000)
print(rf_cat3_all_reduced)

#write.table(importance(rf_cat3_all_reduced), "output/model/importance_rf_cat3_all_reduced.txt", sep="\t")
par(mfrow=c(1,2))
varImpPlot(rf_cat3_all_reduced)
varImpPlot(rf_cat3_all)
par(mfrow=c(1,1))
```


## Random forest classification: ALL control data with 3 classes
```{r rf cat3 all control}
input_cat3_all_control <- data.frame(all_data_control, "tol"=all_info_control$tol_cat3_fve)
levels(input_cat3_all_control$tol)

set.seed(1)
rf_cat3_all_control <- randomForest(tol~. , data=input_cat3_all_control, ntree=1000)
print(rf_cat3_all_control)
```


## Random forest classification: Predict tolerance class for breeder data (Overview)
```{r rf cat3 predict tolerance class for breeder data OVERVIEW}
# ALL data model
rf_cat3_all_predicted_drym_breeder <- predict(rf_cat3_all, all_breeder_data)
confusionMatrix(table(rf_cat3_all_predicted_drym_breeder, all_breeder_info$tol_cat3_fve))
# accuracy: 69.7%

# table(all_breeder_info$cultivar, rf_cat3_all_predicted_drym_breeder)
# table(all_breeder_info$cultivar, all_breeder_info$tol_cat3_fve)

# only CONTROL data model
rf_cat3_all_control_predicted_drym_breeder <- predict(rf_cat3_all_control, all_breeder_data)
confusionMatrix(table(rf_cat3_all_control_predicted_drym_breeder, all_breeder_info$tol_cat3_fve))

# REDUCED model
rf_cat3_all_reduced_predicted_drym_breeder <- predict(rf_cat3_all_reduced, all_breeder_data_vs)
confusionMatrix(table(rf_cat3_all_reduced_predicted_drym_breeder, all_breeder_info$tol_cat3_fve))

# TRAIN data model
confusionMatrix(table(predict(rf_cat3_train, all_breeder_data), all_breeder_info$tol_cat3_fve))$overall[1]
# accuracy: 61.62%

rf_cat3_train_control
```


## Random forest classification: Predict tolerance class for single breeder trials
```{r rf cat3 all predict tolerance class for single breeder trials}
table( predict(rf_cat3_all, all_data_norika_2011), all_info_norika_2011$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_norika_2012), all_info_norika_2012$tol_cat3_fve)

table( predict(rf_cat3_all, all_data_petersgroden_2011), all_info_petersgroden_2011$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_petersgroden_2012), all_info_petersgroden_2012$tol_cat3_fve)

table( predict(rf_cat3_all, all_data_windeby_2011), all_info_windeby_2011$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_windeby_2012), all_info_windeby_2012$tol_cat3_fve)

# years 2011/12
confusionMatrix(table( predict(rf_cat3_all, all_breeder_data_2011), all_breeder_info_2011$tol_cat3_fve))$overall[1]
confusionMatrix(table( predict(rf_cat3_all, all_breeder_data_2012), all_breeder_info_2012$tol_cat3_fve))$overall[1]

# location
table( predict(rf_cat3_all, all_data_norika), all_info_norika$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_petersgroden), all_info_petersgroden$tol_cat3_fve)
table( predict(rf_cat3_all, all_data_windeby), all_info_windeby$tol_cat3_fve)

# accuracy for single trials
confusionMatrix(table( predict(rf_cat3_all, all_data_norika_2011), all_info_norika_2011$tol_cat3_fve))$overall[1]
confusionMatrix(table( predict(rf_cat3_all, all_data_norika_2012), all_info_norika_2012$tol_cat3_fve))$overall[1]

confusionMatrix(table( predict(rf_cat3_all, all_data_petersgroden_2011), all_info_petersgroden_2011$tol_cat3_fve))$overall[1]
confusionMatrix(table( predict(rf_cat3_all, all_data_petersgroden_2012), all_info_petersgroden_2012$tol_cat3_fve))$overall[1]

confusionMatrix(table( predict(rf_cat3_all, all_data_windeby_2011), all_info_windeby_2011$tol_cat3_fve))$overall[1]
confusionMatrix(table( predict(rf_cat3_all, all_data_windeby_2012), all_info_windeby_2012$tol_cat3_fve))$overall[1]
```


# Permutation
```{r permutation}

Y <- all_info$mdrym_fve

set.seed(1234)
A <- replicate(50000, sample(Y))
dim(A)

y.pred <- rf_all$predicted

zufall <- function(y){cor(y.pred, y)} 

perm <- apply(A, 2, zufall)


length(which(perm>=cor(Y, y.pred)))/50000 # 0
1-pnorm(cor(Y,y.pred), mean=mean(perm), sd=sd(perm), lower.tail = TRUE, log.p = FALSE)     #   0
pnorm(cor(Y,y.pred), mean=mean(perm), sd=sd(perm), lower.tail = F, log.p = FALSE)          #   5.341775e-38   

  
mean(perm)    #    -0.0002731715
sd(perm)      #    0.07015797
cor(Y, y.pred) - mean(perm)   #     0.9003524
(cor(Y,y.pred)-mean(perm))/sd(perm)  # 12.83322, dh der echte Wert fuer LOOV_Success. liegt mehr als 26 Standardabweichungen von der Zufallscorr. entfernt

hist(perm, xlim = c(-0.3, 1))
abline(v=cor(Y, y.pred), col="red")

```


# Lasso Model
## Lasso: Define data
```{r lasso define data}
# TEST data
test_x <- as.matrix(test_data)
test_y <- test_info$mdrym_fve
test_y_cat <- test_info$tol_cat3_fve

# use TRAIN data
train_x <- as.matrix(train_data)
train_y <- train_info$mdrym_fve
train_y_cat <- train_info$tol_cat3_fve

# use ALL data
x <- as.matrix(all_data)
y <- all_info$mdrym_fve
y_cat <- all_info$tol_cat3_fve
```


## Lasso: Regression model with ALL data
* default: alpha = 1 --> for lasso penalty (or alpha = 0 for ridge penalty)
* predict needs a value for s, to define the penalty parameter lambda --> here: lambda.min

```{r lasso reg model with all data}
set.seed(1)
lasso_all <- glmnet(x, y)

# Cross-validation
set.seed(1)
lasso_cv_all <- cv.glmnet(x, y)
plot(lasso_cv_all, ylim=c(0, 0.005))

# ridge_cv_all <- cv.glmnet(x, y, alpha=0)
# ridge_cv_all$lambda
# ridge_cv_all$nzero

# use no specific lambda
lasso_all_pred_test <- predict(lasso_all, new = test_x)

# calculate mean error
lasso_all_pred_test_err <- apply( (lasso_all_pred_test - test_y)^2, 2, mean )
points(log(lasso_all$lambda), lasso_all_pred_test_err, col="blue", pch="*")

# use optimal lamba value
lasso_all_pred_all_lambda <- predict(lasso_cv_all, new = x, s="lambda.min")
lasso_all_pred_test_lambda <- predict(lasso_cv_all, new = test_x, s="lambda.min")

# plot observed vs. predicted DRYM
plot(y, lasso_all_pred_all_lambda)
abline(0,1)
cor.test(y, lasso_all_pred_all_lambda)

sqrt(mean((y - lasso_all_pred_all_lambda)^2))

# number of variables in model (for lambda.min)
which(lasso_cv_all$lambda==lasso_cv_all$lambda.min)
# 42
lasso_cv_all$nzero[42]
# 34
```


## Lasso Regression: Predict DRYM for breeder data
```{r lasso reg predict DRYM for breeder data}
# use optimal lamba value
lasso_all_pred_breeder <- predict(lasso_cv_all, new = as.matrix(all_breeder_data), s = "lambda.min")
lasso_all_pred_breeder_2011 <- predict(lasso_cv_all, new = as.matrix(all_breeder_data_2011), s = "lambda.min")
lasso_all_pred_breeder_2012 <- predict(lasso_cv_all, new = as.matrix(all_breeder_data_2012), s = "lambda.min")

plot(all_breeder_info$mdrym_fve, lasso_all_pred_breeder, 
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19,
     xlim = c(-0.18, 0.15), ylim = c(-0.18, 0.15))
abline(0, 1, col = "red")

cor.test(all_breeder_info$mdrym_fve, lasso_all_pred_breeder)
sqrt(mean((all_breeder_info$mdrym_fve - lasso_all_pred_breeder)^2))

# 2011/12
cor.test(all_breeder_info_2011$mdrym_fve, lasso_all_pred_breeder_2011)$estimate # 0.55
cor.test(all_breeder_info_2012$mdrym_fve, lasso_all_pred_breeder_2012)$estimate # 0.66

# single breeder trials
lasso_all_pred_norika_2011 <- predict(lasso_cv_all, new = as.matrix(all_data_norika_2011), s = "lambda.min")
lasso_all_pred_norika_2012 <- predict(lasso_cv_all, new = as.matrix(all_data_norika_2012), s = "lambda.min")
lasso_all_pred_petersgroden_2011 <- predict(lasso_cv_all, new = as.matrix(all_data_petersgroden_2011), s = "lambda.min")
lasso_all_pred_petersgroden_2012 <- predict(lasso_cv_all, new = as.matrix(all_data_petersgroden_2012), s = "lambda.min")
lasso_all_pred_windeby_2011 <- predict(lasso_cv_all, new = as.matrix(all_data_windeby_2011), s = "lambda.min")
lasso_all_pred_windeby_2012 <- predict(lasso_cv_all, new = as.matrix(all_data_windeby_2012), s = "lambda.min")

# correlations
cor(lasso_all_pred_norika_2011, all_info_norika_2011$mdrym_fve)
cor(lasso_all_pred_norika_2012, all_info_norika_2012$mdrym_fve)
cor(lasso_all_pred_petersgroden_2011, all_info_petersgroden_2011$mdrym_fve)
cor(lasso_all_pred_petersgroden_2012, all_info_petersgroden_2012$mdrym_fve)
cor(lasso_all_pred_windeby_2011, all_info_windeby_2011$mdrym_fve)
cor(lasso_all_pred_windeby_2012, all_info_windeby_2012$mdrym_fve)
```


### Lasso Regression: Comparison to Random Forest Norika
```{r lasso reg comparison to random forest Norika}
pdf("figures/model/prediction_rf_lasso_norika.pdf", width=10, height=5)
par(mfrow=c(1,2))
# Norika 2011
plot(all_info_norika_2011$mdrym_fve, predict(rf_all, all_data_norika_2011), col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "Random Forest")
text(all_info_norika_2011$mdrym_fve, predict(rf_all, all_data_norika_2011), 
     labels = all_info_norika_2011$cultivar, cex = 0.6)
abline(0, 1, col = "red")

plot(all_info_norika_2011$mdrym_fve, lasso_all_pred_norika_2011, col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "LASSO")
text(all_info_norika_2011$mdrym_fve, lasso_all_pred_norika_2011, 
     labels = all_info_norika_2011$cultivar, cex = 0.6)
abline(0, 1, col = "red")

# Norika 2012
plot(all_info_norika_2012$mdrym_fve, predict(rf_all, all_data_norika_2012), col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "Random Forest")
text(all_info_norika_2012$mdrym_fve, predict(rf_all, all_data_norika_2012), 
     labels = all_info_norika_2012$cultivar, cex = 0.6)
abline(0, 1, col = "red")

plot(all_info_norika_2012$mdrym_fve, lasso_all_pred_norika_2012, col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "LASSO")
text(all_info_norika_2012$mdrym_fve, lasso_all_pred_norika_2012, 
     labels = all_info_norika_2012$cultivar, cex = 0.6)
abline(0, 1, col = "red")
dev.off()
```


### Lasso Regression: Comparison to Random Forest Petersgroden
```{r lasso reg comparison to random forest Petersgroden}
pdf("figures/model/prediction_rf_lasso_petersgroden.pdf", width=10, height=5)
par(mfrow=c(1,2))
# Petersgroden 2011
plot(all_info_petersgroden_2011$mdrym_fve, predict(rf_all, all_data_petersgroden_2011), col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "Random Forest")
text(all_info_petersgroden_2011$mdrym_fve, predict(rf_all, all_data_petersgroden_2011), 
     labels = all_info_petersgroden_2011$cultivar, cex = 0.6)
abline(0, 1, col = "red")

plot(all_info_petersgroden_2011$mdrym_fve, lasso_all_pred_petersgroden_2011, col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "LASSO")
text(all_info_petersgroden_2011$mdrym_fve, lasso_all_pred_petersgroden_2011, 
     labels = all_info_petersgroden_2011$cultivar, cex = 0.6)
abline(0, 1, col = "red")

# Petersgroden 2012
plot(all_info_petersgroden_2012$mdrym_fve, predict(rf_all, all_data_petersgroden_2012), col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "Random Forest")
text(all_info_petersgroden_2012$mdrym_fve, predict(rf_all, all_data_petersgroden_2012), 
     labels = all_info_petersgroden_2012$cultivar, cex = 0.6)
abline(0, 1, col = "red")

plot(all_info_petersgroden_2012$mdrym_fve, lasso_all_pred_petersgroden_2012, col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "LASSO")
text(all_info_petersgroden_2012$mdrym_fve, lasso_all_pred_petersgroden_2012, 
     labels = all_info_petersgroden_2012$cultivar, cex = 0.6)
abline(0, 1, col = "red")
dev.off()
```


### Lasso Regression: Comparison to Random Forest Windeby
```{r lasso reg comparison to random forest Windeby}
pdf("figures/model/prediction_rf_lasso_windeby.pdf", width=10, height=5)
par(mfrow=c(1,2))
# Windeby 2011
plot(all_info_windeby_2011$mdrym_fve, predict(rf_all, all_data_windeby_2011), col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "Random Forest")
text(all_info_windeby_2011$mdrym_fve, predict(rf_all, all_data_windeby_2011), 
     labels = all_info_windeby_2011$cultivar, cex = 0.6)
abline(0, 1, col = "red")

plot(all_info_windeby_2011$mdrym_fve, lasso_all_pred_windeby_2011, col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "LASSO")
text(all_info_windeby_2011$mdrym_fve, lasso_all_pred_windeby_2011, 
     labels = all_info_windeby_2011$cultivar, cex = 0.6)
abline(0, 1, col = "red")

# Windeby 2012
plot(all_info_windeby_2012$mdrym_fve, predict(rf_all, all_data_windeby_2012), col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "Random Forest")
text(all_info_windeby_2012$mdrym_fve, predict(rf_all, all_data_windeby_2012), 
     labels = all_info_windeby_2012$cultivar, cex = 0.6)
abline(0, 1, col = "red")

plot(all_info_windeby_2012$mdrym_fve, lasso_all_pred_windeby_2012, col="lightgrey",
     xlab = "observed DRYM", ylab = "predicted DRYM", pch = 19, main = "LASSO")
text(all_info_windeby_2012$mdrym_fve, lasso_all_pred_windeby_2012, 
     labels = all_info_windeby_2012$cultivar, cex = 0.6)
abline(0, 1, col = "red")
dev.off()

par(mfrow=c(1,1))
```


## Lasso: Multinomial model with TRAINING data
A special option for multinomial regression is ``type.multinomial``, which allows the usage of a grouped lasso penalty if ``type.multinomial = "grouped"``. 
This will ensure that the multinomial coefficients for a variable are all in or out together, just like for the multi-response Gaussian.
```{r lasso multinomial model with train data}
set.seed(1)
lasso_cat_train <- glmnet(train_x, train_y_cat, family = "multinomial")

# Cross-validation
set.seed(1)
lasso_cat_cv_train <- cv.glmnet(train_x, train_y_cat, family = "multinomial")
plot(lasso_cat_cv_train)

which(lasso_cat_cv_train$lambda == lasso_cat_cv_train$lambda.min)
# 46
lasso_cat_cv_train$nzero[46]
# 15

# use optimal lamba value
lasso_cat_train_pred_tain_lambda <- predict(lasso_cat_cv_train, new = train_x, s = "lambda.min", type = "class")
lasso_cat_train_pred_test_lambda <- predict(lasso_cat_cv_train, new = test_x, s = "lambda.min", type = "class")

# table observed vs. predicted DRYM
table(train_y_cat, lasso_cat_train_pred_tain_lambda)
table(test_y_cat, lasso_cat_train_pred_test_lambda)
```


## Lasso: Multinomial model with ALL data
```{r lasso multinomial model with all data}
set.seed(1)
lasso_cat_all <- glmnet(x, y_cat, family = "multinomial")

# Cross-validation
set.seed(1)
lasso_cat_cv_all <- cv.glmnet(x, y_cat, family = "multinomial")
plot(lasso_cat_cv_all)
which(lasso_cat_cv_all$lambda == lasso_cat_cv_all$lambda.min)
# 53
lasso_cat_cv_all$nzero[53]
# 17

# use optimal lamba value
lasso_cat_all_pred_all_lambda <- predict(lasso_cat_cv_all, new = x, s = "lambda.min", type = "class")

# plot observed vs. predicted DRYM
table(y_cat, lasso_cat_all_pred_all_lambda)
```


## Lasso: Multinomial model: Predict DRYM for breeder data
```{r lasso multinomial predict DRYM for breeder data}

func_pred_new_values <- function(new_x, obs_y, pred_model){
  
  pred_y <- predict(pred_model, new = as.matrix(new_x), 
                    s = "lambda.min", type  = "class")
  # convert to factor
  pred_y_f <- factor(pred_y[,1], levels = c("low", "mid", "high"))
  
  # print(table(pred_y_f, obs_y))
  print(confusionMatrix(table(pred_y_f, obs_y)))
}

# all breeder data
func_pred_new_values(new_x = all_breeder_data,
                     obs_y = all_breeder_info$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

# Norika 2011/12
func_pred_new_values(new_x = all_data_norika_2011,
                     obs_y = all_info_norika_2011$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

func_pred_new_values(new_x = all_data_norika_2012,
                     obs_y = all_info_norika_2012$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

# Petersgroden 2011/12
func_pred_new_values(new_x = all_data_petersgroden_2011,
                     obs_y = all_info_petersgroden_2011$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

func_pred_new_values(new_x = all_data_petersgroden_2012,
                     obs_y = all_info_petersgroden_2012$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

# Windeby 2011/12
func_pred_new_values(new_x = all_data_windeby_2011,
                     obs_y = all_info_windeby_2011$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)

func_pred_new_values(new_x = all_data_windeby_2012,
                     obs_y = all_info_windeby_2012$tol_cat3_fve,
                     pred_model = lasso_cat_cv_all)
```



# PLS-DA (by mixomics)
```{r PLS-DA}
# use ALL data
x <- as.matrix(all_data)
y_cat <- all_info$tol_cat3_fve
class(y_cat)

plsda_perf_all <- mixOmics::plsda(x, y_cat, ncomp = 10)
perf_plsda <- mixOmics::perf(plsda_perf_all, validation = 'Mfold', folds = 5,
                   progressBar = TRUE, nrepeat = 10)

# perf_plsda <- perf(plsda_perf_all, validation = 'loo',
#                    progressBar = TRUE, nrepeat = 10)

perf_plsda$error.rate
plot(perf_plsda, overlay = 'measure', sd=TRUE)
# choose 5 comps

plsda_perf_all <- mixOmics::plsda(x, y_cat, ncomp = 5)

plotIndiv(plsda_perf_all , comp = c(1,2),
          group = y_cat, ind.names = FALSE, 
          ellipse = TRUE, legend = TRUE, title = 'PLSDA comp 1 - 2')

plotIndiv(plsda_perf_all , comp = c(2,3),
          group = y_cat, ind.names = FALSE, 
          ellipse = TRUE, legend = TRUE, title = 'PLSDA comp 1 - 2')

```


# Save workspace and sessionInfo
```{r save workspace}
save.image("qpcr_model_field_breeder.RData")
sessionInfo()
```

